---
title: "temp"
output: html_document
date: "2025-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Finding optimal parameters using model 7

#The goal is to run essays that result in a 50% binding.
```{r Find optimal parameters model 7}

# choose one model (m7: buffer + weight + spike)
coef <- coef(m7)

# Set target binding
target_binding <- 50

# FUNCTION to Solve for weight, assuming spike = 0
# 50% - intercept - (buffer1 * 1) - (spike * 0) / weight  

solve_for_weight <- function(dilution_value, spike_value = 0) {
  (target_binding - coef[1] - coef[3] * dilution_value - coef[4] * spike_value) / coef[2]
}

# Find the weight that gives 50% binding whenspike is 0
# dilution = 250
optimal_weight <- solve_for_weight(dilution_value = 1)
optimal_weight
# dilution = 60
optimal_weight <- solve_for_weight(dilution_value = 0)
optimal_weight

# Find the weight that gives 50% binding when spike is 1
# dilution = 250
optimal_weight <- solve_for_weight(dilution_value = 1, spike_value = 1)
optimal_weight
# dilution = 60
optimal_weight <- solve_for_weight(dilution_value = 0, spike_value = 1)
optimal_weight

```


# visualize results

```{r}

## Buffer = 1, Spike = 0
new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(1), spike = as.factor(0))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 60, spike = 1)",
       x = "Weight",
       y = "Predicted Binding Percentage")

## Buffer = 1, Spike = 1
new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(1), spike = as.factor(1))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 250, spike = 1)",
       x = "Weight",
       y = "Predicted Binding Percentage")


## Buffer = 0, Spike = 1

new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(0), spike = as.factor(1))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 250, spike = 0)",
       x = "Weight",
       y = "Predicted Binding Percentage")

## Buffer = 0, Spike = 0

new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(0), spike = as.factor(0))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 60, spike = 0)",
       x = "Weight",
       y = "Predicted Binding Percentage")



# FUNCTION to Solve for weight, assuming spike = 0
# 50% - intercept - (buffer1 * 1) - (spike * 0) / weight  

solve_for_weight <- function(dilution_value, spike_value = 0) {
  (target_binding - coef[1] - coef[3] * dilution_value - coef[4] * spike_value) / coef[2]
}


# Loop over different dilution values to find the optimal weight for 50% binding
# here, dilution 0 means 60uL, and 1 is 250 uL
for (buffer in seq(0, 1, by = 0.1)) {
  optimal_weight <- solve_for_weight(buffer)
  cat("Dilution:", buffer, "-> Optimal Weight for 50% Binding:", optimal_weight, "\n")
}
```


# visualize results
## Buffer = 1, Spike = 0
new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(1), spike = as.factor(0))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 60, spike = 1)",
       x = "Weight",
       y = "Predicted Binding Percentage")

## Buffer = 1, Spike = 1
new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(1), spike = as.factor(1))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 250, spike = 1)",
       x = "Weight",
       y = "Predicted Binding Percentage")


## Buffer = 0, Spike = 1

new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(0), spike = as.factor(1))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 250, spike = 0)",
       x = "Weight",
       y = "Predicted Binding Percentage")

## Buffer = 0, Spike = 0

new_data <- expand.grid(weight = seq(min(weight), max(weight), length.out = 100),
                        buffer = as.factor(0), spike = as.factor(0))  # Set buffer and spike to a fixed value for simplicity

# Predict the binding percentage for the new data
new_data$predicted_binding <- predict(m7, newdata = new_data)

# Plot the predicted binding percentage against weight
ggplot(new_data, aes(x = weight, y = predicted_binding)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +  # Highlight 50% binding
  labs(title = "Predicted Binding Percentage vs Weight (dilution = 60, spike = 0)",
       x = "Weight",
       y = "Predicted Binding Percentage")



# FUNCTION to Solve for weight, assuming spike = 0
# 50% - intercept - (buffer1 * 1) - (spike * 0) / weight  

solve_for_weight <- function(dilution_value, spike_value = 0) {
  (target_binding - coef[1] - coef[3] * dilution_value - coef[4] * spike_value) / coef[2]
}


# Loop over different dilution values to find the optimal weight for 50% binding
# here, dilution 0 means 60uL, and 1 is 250 uL
for (buffer in seq(0, 1, by = 0.1)) {
  optimal_weight <- solve_for_weight(buffer)
  cat("Dilution:", buffer, "-> Optimal Weight for 50% Binding:", optimal_weight, "\n")
}

```



Work in progress 
after this line
______________________
```{r echo = FALSE}




```

```{r echo = FALSE}



#wflow_publish("./analysis/ELISA_visualizations.Rmd")

#wflow_status()


```

