---
title: "Analysis of final values - test7"
author: "Paloma"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---
# Introduction

Here I plan to include measurements and visualizations of the accuracy, precision, and linearity of ELISA #6. I will do this using two different ways of calculating final values: the standard method not accounting for spike, and an alternative method accounting for it. The way the spike is included in the formula is defined in file ELISA_Calc_FinalVals_test6.html.

*Plate description*

Category	  Description
---------  ---------------------------------------------------------------
T        	Non-spiked 


```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries, echo = FALSE, warning = FALSE, message=FALSE}
# Install libraries

library(knitr)
library(RColorBrewer)
library(stats)
library(coefplot)
library(arm)
library(bbmle)
library(plyr)
library(ggpmisc)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
```

```{r loading}
# Loading data
std <- read.csv("./data/Test6/standard_data_test6.csv")
cort_vals<- read.csv("./data/Test6/Data_cort_values_method_ALL.csv")
```

# Overall Results
## A) Standard method

```{r loading valsA}
# remove NSB
data <- cort_vals
```

```{r expABC.Final_pg.mg_A, fig.width=8, fig.height=6}
# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_pg.mg_A)) 
a$Sample_num <- 1:length(a$Sample)
a$Failed_samples[is.na(a$Failed_samples)] <- "OK"
cat(paste("Data has", nrow(a), "data points", sep= " "))


## PLOT WITHOUT POOR QUALITY SAMPLES ### 
exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "A2"))

ggplot(exp, aes(x = Sample_num, 
                 y = Final_pg.mg_A, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.2, aes(group = exp$Category)) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.3) + 
 stat_poly_eq(aes(x = Sample_num, 
                  label = paste(..rr.label.., sep = "~")),
                  formula = y ~ x, parse = TRUE, size = 2) +
  labs(title = "Final vals (A) not accounting for spike",
       subtitle = "removed high CV samples",
       x = "Sample", y = "Cortisol concentration pg/mg, Method A") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

## B) Accounting for spike (Nist et al. 2020)

```{r expABC.Final_pg.mg_B, fig.width=8, fig.height=6, }

## PLOT WITHOUT POOR QUALITY SAMPLES ### 

ggplot(exp, aes(x = Sample_num, 
                 y = Final_pg.mg_B, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.2) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.5) + 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  labs(title = "Final vals (B) accounting for spike",
       subtitle = "removed high CV samples",
       x = "Sample", y = "Cortisol concentration pg/mg, Method B") +
  scale_y_continuous(n.breaks = 10)  +
  theme_minimal()

```

## C) Sam's Method
```{r expABC.Final_pg.mg_C}

## PLOT WITHOUT POOR QUALITY SAMPLES ### 

ggplot(exp, aes(x = Sample_num, 
                 y = Final_pg.mg_C, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.2) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  coord_cartesian(ylim = c(0, 55)) + 
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.5) + 
  labs(title = "Final values, method C",
       subtitle = "Removed low quality samples",
       x = "Sample", y = "Cortisol concentration pg/mg, Method C") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

```{r ABC.Final_pg.mg_C-all, echo = FALSE}
## PLOT ALL SAMPLES ### 

ggplot(a, aes(x = Sample_num, 
                 y = Final_pg.mg_C, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.2) +
  geom_point(size = 2) +
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.5) + 
  coord_cartesian(ylim = c(0, 55)) + 
  labs(title = "Final values, method C",
       subtitle = "Included all samples",
       x = "Sample", y = "Cortisol concentration pg/mg, Method C") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```


# Comparing all calculation methods
## Summary and boxplots
```{r 4-results-boxplot, echo = FALSE}
# Save result
cat("Results:")
tail(data[,c("Sample", "Final_pg.mg_A", "Final_pg.mg_B", "Final_pg.mg_C")])
data4<-data

# View summary
cat("Summary of final values for each calculation method")
cat("Method A:")
summary(data$Final_pg.mg_A)
cat("Method B:")
summary(data$Final_pg.mg_B)
cat("Method C:")
summary(data$Final_pg.mg_C)

# Reshape from wide to long format
data_long <- data %>%
  dplyr::select(Final_pg.mg_A, Final_pg.mg_B, Final_pg.mg_C) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Method",
    values_to = "pg_mg"
  )

# clean group names for x-axis labels
data_long$Group <- gsub("Final_pg.mg_", "", data_long$Method)

# Plot boxplots
ggplot(data_long, aes(x = Group, y = pg_mg, fill = Method)) +
  geom_boxplot(outlier.shape = 21, outlier.fill = "white", outlier.color = "black", width = 0.15, alpha = 0.4) +
  geom_violin(alpha = 0.25) +
  labs(
    title = "Test4: Final values using calculation methods A, B, and C",
    x = "Method",
    y = "pg/mg"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

## Changes in each data point

### ABC calculation methods
```{r 4-results-scatterplot, echo = FALSE}

# Reshape data for plotting
data_long <- data %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_B, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

ggplot(data_long, aes(x = Sample, y = Final_pg.mg, color = Method, group = Method)) +
  geom_line(aes(group = Sample), color = "gray75", linewidth = 0.5) +
  geom_point(size = 2.8, alpha = 0.7, position = position_dodge(width=0.5)) +
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue",
               "Final_pg.mg_B" = "orange",
               "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A (Traditional)",
               "Method B (Nist et al.)",
               "Method C (Sam's)")
  ) +
  labs(
    title = "Hair Cortisol by Sample and Calculation Method",
    x = "Sample ID",
    y = "Final Cortisol (pg/mg)",
    color = "Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "top"
  )


```

### A and C calculation methods
```{r 4-results-A and C, echo =FALSE}

# Reshape to long format
data_long <- data %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

# Plot with lines connecting A and C values for each sample
ggplot(data_long, aes(x = Sample, y = Final_pg.mg, color = Method, group = Method)) +
  geom_line(aes(group = Sample), linewidth = 0.6, color = "gray70") +  # light line connecting methods
  geom_point(size = 2.8, alpha = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue", "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A", "Method C")
  ) +
  labs(
    title = "Comparing final values obtained using 2 methods, by Sample",
    x = "Sample ID",
    y = "Final Cortisol Concentration (pg/mg)",
    color = "Calculation Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "top"
  )

```
### A and C, good quality only

```{r}
# Filter to include only good quality samples
data_ok <- data %>%
    filter(!is.na(Binding.Perc)) %>%
  filter(Failed_samples == "OK")

# Reshape data for plotting
data_long_ok <- data_ok %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

head(data_long_ok)
# Plot for OK samples only
ggplot(data_long_ok, aes(x = Sample, y = Final_pg.mg, color = Method)) +
  geom_line(aes(group = Sample), color = "gray75", size = 0.5) +
  geom_point(size = 3, alpha = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue",
               "Final_pg.mg_B" = "orange",
               "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A (Basic)",
               "Method B (Subtraction + x2)",
               "Method D (Spike contribution)")
  ) +
  labs(
    title = "Hair Cortisol (Good Samples Only) — Test4",
    x = "Sample ID",
    y = "Final Cortisol (pg/mg)",
    color = "Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "top"
  )
```

# Precision

## Method A (Standard)
```{r expE-concvals-final-vals-A}

# Scatter plot of Experiment a, observed binding % 
a <- filter(data, Category == "P", !is.na(Final_pg.mg_A)) 
a$Sample_num <- 1:length(a$Sample)

exp <- a %>%
  mutate(Group = substr(Sample, 1, 3))

cv_values <- exp %>% 
  group_by(Group) %>%
  dplyr::summarise(
    cv_cort = round((sd(Final_pg.mg_A) / mean(Final_pg.mg_A)) * 100, 1), 
    mean_cort = round( mean( Final_pg.mg_A, na.rm = TRUE), 1),
    sd_cort = round(sd( Final_pg.mg_A, na.rm = TRUE), 1)
  )
cat("Variation in final cortisol values using A (good quality data only):")
print(cv_values)


ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_pg.mg_A)) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_cort,
                     ymin = mean_cort - sd_cort, 
                     ymax = mean_cort + sd_cort),
                color = "orange", size = 0.5, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
   labs(title = "Experiment E: Final conc vals (methodA) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_pg.mg_A") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 5.5, y = 65, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 8.5, y = 65, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 11.5, y = 65, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 5.5, y = 60, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 8.5, y = 60, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 11.5, y = 60, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP3"],2)), size = 4) 

```

## Method B (Nist et al)
```{r expE-concvals-final-vals-B}

# Scatter plot of Experiment a, observed binding % 
cv_values <- exp %>% 
  group_by(Group) %>%
  dplyr::summarise(
    CV_cort = round((sd( Final_pg.mg_B) / mean(Final_pg.mg_B)) * 100, 1), 
    mean_cort = mean( Final_pg.mg_B, na.rm = TRUE),
    sd_cort = sd( Final_pg.mg_B, na.rm = TRUE)
  )
    
print(cv_values)

ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_pg.mg_B)) +
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_cort,
                      ymin = mean_cort - sd_cort, ymax = mean_cort + sd_cort),
                  color = "orange", size = 0.5, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.8) +  
   labs(title = "Experiment E: Final conc vals (methodB) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_pg.mg_B") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$CV_cort[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$CV_cort[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$CV_cort[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP3"],2)), size = 4) 

```


## Method C (Sam's)
```{r expE-concvals-final-vals-c}

# Scatter plot of Experiment C, observed binding % 
cv_values <- exp %>% 
  group_by(Group) %>%
  dplyr::summarise(
    mean_cort = round( mean( Final_pg.mg_C, na.rm = TRUE), 1),
    sd_cort = round(sd( Final_pg.mg_C, na.rm = TRUE), 1),
    cv_cort = round((sd( Final_pg.mg_C ) / mean( Final_pg.mg_C)) * 100, 1), 
  )

print(cv_values)
ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_pg.mg_C)) +
geom_pointrange(data = cv_values,
                 aes(x = c(6,9,12), 
                     y = mean_cort,
                     ymin = mean_cort - sd_cort, 
                     ymax = mean_cort + sd_cort),
                 color = "orange", 
                 size = 0.5, alpha = 0.7) +
  geom_point(size = 3, 
           alpha = 1, 
           aes(fill = factor(Failed_samples),
               color = factor(Failed_samples))) +
  geom_smooth(method = "lm", se = TRUE, 
              linewidth = 0.8, 
              alpha = 0.15, 
              color = "orange") +
  geom_text(aes(label = Sample), 
            size = 3, vjust = -1, 
            hjust = -0.5, 
            color = "gray30") +
   labs(title = "Experiment E: Final conc vals (methodC) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_pg.mg_C") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 38, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 38, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 38, label = paste0("CV: ", cv_values$cv_cort[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_cort[cv_values$Group == "TP3"],2)), size = 4) 

```


# Accuracy (using mean for A + spike contribution)

Accuracy, or recovery %, is the ratio between observed and expected concentration, multiplied by 100. 
Expected values are defined using the dilution factor, starting with the concentration value of the neat sample. 

- 80–120% is generally acceptable for biological assays.

- 70–130% may be acceptable at very low concentrations.

**Problem**: How should I define expected vals for final cortisol values? Can this be done only using binding percentage?

# Parallelism

```{r}
# Load the dataset
df <- data4

# Clean data 
df <- df %>% filter(!is.na(Final_pg.mg_C), !Sample %in% c("NSB", "Blank", "B0"))
df$Category <- as.factor(df$Category)

# 1. === PARALLELISM TEST ===
# Plot concentration vs. dilution to check for proportional decrease

ggplot(df, aes(x = log10(as.numeric(Dilution_sample)), y = Final_pg.mg_C , color = Category)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.05) +
  labs(title = "Parallelism Test, final cortisol value", 
       x = "Log10(Dilution_sample)", 
       y = "Final Cortisol (pg/mg)") + 
  theme_minimal()

# Optional: Run ANOVA for parallelism by slope comparison
parallel_model <- lm(Final_pg.mg_C  ~ log10(Dilution_sample) * Category, data = df)
summary(parallel_model)
```

## Slope
|	Term |	Estimate |		p-value |		Interpretation |	
|-----|--------|	--------- |	----------- |	
 |	(Intercept)	 |	35.08	 |	1.7e-05	 |	Baseline value for TA at no dilution |	
 |	CategoryC |		–19.89 |		0.0473 *	 |	TC starts significantly lower than TA |	
 |	CategoryD |		–18.70 |		0.0612 .	 |	TD also starts lower, marginally significant |	
 |	CategoryP	 |	–15.34 |		0.0517 . |		TP slightly lower, borderline significant |	
 |	Interaction (log10 × CategoryB)	 |	+17.33	 |	0.0598 .	 |	TB's slope is steeper than TA (marginal) |	

TC and TD show parallel dilution curves relative to TA, suggesting you can safely apply Method D.

TB may have a steeper slope, suggesting spiking only the first tube might amplify dilution effect slightly.

## Binding
```{r plot parallelism_binding}
ggplot(df, aes(x = log10(as.numeric(Dilution_sample)), 
               y = Binding.Perc, color = Category)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Parallelism Test (Binding)", x = "Log10(Dilution_sample)", y = "Binding %")

# Optional: Run ANOVA for parallelism by slope comparison
parallel_model <- lm(Binding.Perc  ~ log10(Dilution_sample) * Category, data = df)
summary(parallel_model)

```

|Term |		Estimate|	p-value	|Interpretation|
|------|-----------|---------|------------|
|(Intercept)|	11.70|	0.0081 **	| Baseline binding % for Category A (TA) at no dilution|
|log10(Dilution_sample)|	49.88	|1.6e-13 ***|	Strong increase in binding % with increasing dilution|
|CategoryB|	4.71	|0.424	|TB baseline slightly higher than TA, not significant|
|CategoryC|	10.10	|0.093 .|	TC baseline marginally higher than TA|
|CategoryD|	6.19	|0.295	|TD not significantly different from TA at baseline|
|CategoryP|	7.94	|0.093 .|	TP marginally higher binding at baseline|

log10(Dilution_sample):CategoryB	–4.66	0.391	TB slope not significantly different from TA
log10(Dilution_sample):CategoryC	–3.19	0.556	TC slope ≈ TA
log10(Dilution_sample):CategoryD	–0.36	0.947	TD slope ≈ TA 

 Strong Main Effect of Dilution
Binding % increases significantly with higher dilution (p < 0.001).

This is expected: lower concentration = higher % binding in competitive ELISA.

2. No Significant Interaction Terms
The interaction terms (slope differences) are not significant (all p > 0.39).

This means the binding curves across all categories are parallel.

3. Intercept Differences (Baseline Shift)
Some categories (notably TC and TP) start with higher binding %, but this is only marginally significant (p ≈ 0.09).

This suggests potential small matrix effects, but not enough to invalidate comparison.

```{r}
# Log buffer as dilution proxy
df$log_dilution <- log10(df$Dilution_sample)

# Parallelism test for TA vs TC
df_sub <- df %>% filter(Category %in% c("A", "C"))

model_parallel <- lm(Final_pg.mg_C ~ log_dilution * Category, data = df_sub)
summary(model_parallel)

```

|Term|	Estimate|	p-value	| Interpretation |
|----|---------|--------|---------------|
|(Intercept)	|35.08	|0.0003	|Cortisol for TA at dilution = 1 (log10 = 0)|
|log_dilution	|1.58	|0.794	|Slope of dilution for TA — not significant|
|CategoryC|	–19.89	|0.052 .|	TC starts ~20 pg/mg lower than TA — borderline significant|
|Interaction (log_dilution:CategoryC)|	9.44|	0.282	|TC slope vs TA — not significant → ✅ parallel |slopes|

🔍 Parallelism holds: no significant difference in slope between TA and TC (p = 0.282)

⚠️ Intercept difference between TA and TC is borderline significant (p = 0.052)

TC starts lower than TA, suggesting a shift in baseline cortisol, but dilution behavior is similar


```{r}

# Filter dataset to include only TA (A) and TC (C)
df_sub <- df %>%
  filter(Category %in% c("A", "C")) %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "A2")) %>%
  mutate(
    Category = factor(Category, levels = c("A", "C")),
    log_dilution = log10(Dilution_sample)
  )


# Plot
ggplot(df_sub, aes(x = log_dilution, y = Final_pg.mg_C, color = Category)) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 1.2, alpha = 0.3) +
  geom_point(size = 3, alpha = 0.8) +

  labs(
    title = "Parallelism Test: TA vs TC",
    subtitle = "Final Cortisol (pg/mg) vs Log10(Dilution) with 95% CI",
    x = "Log10(Dilution Sample)",
    y = "Final Cortisol (pg/mg)",
    color = "Category"
  ) +
  scale_color_manual(
    values = c("A" = "steelblue", "C" = "darkgreen"),
    labels = c("TA (Non-Spiked)", "TC (All Dilutions Spiked)")
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  )


```
If slope ≈ 0 and variance is low → good precision.

```{r}

# Load your data

# Extract meaningful category label
df$Category <- str_extract(df$Sample, "^T[ABCD]")

# Keep only A–D dilution categories
df <- df %>% filter(Category %in% c("TA", "TB", "TC", "TD"))

# Create log dilution
df$log_dilution <- log10(df$Dilution_sample)

# All pairwise category combinations
pairs <- combn(unique(df$Category), 2, simplify = FALSE)

# Loop and test each pair for both variables
for (pair in pairs) {
  sub_df <- df %>% filter(Category %in% pair)
  pair_label <- paste(pair, collapse = "_vs_")
  
  ### 1. Final Values Model
  model_final <- lm(Final_pg.mg_C ~ log_dilution * Category, data = sub_df)
  p_final <- tidy(model_final) %>% 
    filter(str_detect(term, "log_dilution:Category")) %>%
    pull(p.value)
  
  ### 2. Binding % Model
  model_binding <- lm(Binding.Perc ~ log_dilution * Category, data = sub_df)
  p_bind <- tidy(model_binding) %>% 
    filter(str_detect(term, "log_dilution:Category")) %>%
    pull(p.value)
  
  ### Plot for Final Values
p1 <- ggplot(sub_df, aes(x = log_dilution, y = Final_pg.mg_C, color = Category)) +
   geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +  
  geom_point(size = 2) +
   
    labs(
      title = paste("Final Values –", pair_label),
      subtitle = paste("Interaction p =", round(p_final, 4)),
      x = "log10(Dilution)", y = "Final Cortisol (pg/mg)"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
  print(p1)
  
  ### Plot for Binding %
p2 <- ggplot(sub_df, aes(x = log_dilution, y = Binding.Perc, color = Category)) +
   geom_smooth(method = "lm", se = TRUE, alpha = 0.3) + 
   geom_point(size = 2) +

    labs(
      title = paste("Binding % –", pair_label),
      subtitle = paste("Interaction p =", round(p_bind, 4)),
      x = "log10(Dilution)", y = "Binding (%)"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
  print(p2)
  
  ### Print results to console
  cat("\n=== ", pair_label, " ===\n")
  cat("Interaction p (Final values):", round(p_final, 4), "\n")
  cat("Interaction p (Binding %):   ", round(p_bind, 4), "\n")
  }

```
|Comparison	|Final Values p|	Binding % p|	Interpretation|
|--------|--------|--------|-------|
|TA vs TB|	0.164	|0.381|	Not significantly different slopes. Likely parallel|
|TA vs TC	|0.366|	0.604	|Parallel; spiking all tubes in TC doesn't distort slope much.|
|TA vs TD|	0.951|	0.928	|Extremely parallel — almost identical slopes.|
|TB vs TC|	0.479|	0.845	|No significant slope difference. Spike methods act similarly.|
|TB vs TD|	0.162	|0.471	|Slight trend in slope difference, but still not statistically significant.|
|TC vs TD|	0.367	|0.674	|Parallel. Spiking either all tubes (TC) or just first tube (TD) gives similar dilution response.|

All p-values are > 0.05, meaning there is no strong evidence of non-parallelism.

This suggests serial dilutions are behaving consistently, even with different spiking strategies.

The fact that binding percentage curves are also parallel confirms this is not an artifact of back-calculation — it's real behavior at the assay level.


```{r test4}
# === Load and prepare data ===
df <- data4 %>%
  mutate(Category = str_extract(Sample, "^T[ABCD]")) %>%
  filter(Category %in% c("TA", "TB", "TC", "TD")) %>%
  mutate(log_dilution = log10(Dilution_sample))

# === Updated and safe function ===
extract_intercepts_with_pval <- function(model, response_label) {
  model_tidy <- tidy(model, conf.int = TRUE)
  intercept_ta <- model_tidy %>% filter(term == "(Intercept)") %>% pull(estimate)

  intercepts <- model_tidy %>%
    filter(term == "(Intercept)" | str_detect(term, "Category")) %>%
    mutate(
      Category = case_when(
        term == "(Intercept)" ~ "TA",
        str_detect(term, "TB") ~ "TB",
        str_detect(term, "TC") ~ "TC",
        str_detect(term, "TD") ~ "TD"
      ),
      Estimate = if_else(Category == "TA", estimate, intercept_ta + estimate),
      CI_low = if_else(Category == "TA", conf.low, conf.low + intercept_ta),
      CI_high = if_else(Category == "TA", conf.high, conf.high + intercept_ta),
      P_Value = if_else(Category == "TA", NA_real_, p.value),
      Variable = response_label
    ) %>%
    dplyr::select(Variable, Category, Estimate, CI_low, CI_high, P_Value)

  return(intercepts)
}

# === Fit models ===
mod_final <- lm(Final_pg.mg_C ~ log_dilution + Category, data = df)
mod_binding <- lm(Binding.Perc ~ log_dilution + Category, data = df)

# === Extract intercepts ===
intercepts_final <- extract_intercepts_with_pval(mod_final, "Final Cortisol (pg/mg)")
intercepts_binding <- extract_intercepts_with_pval(mod_binding, "Binding Percentage")

# === Combine and reshape for table ===
intercepts_all <- bind_rows(intercepts_final, intercepts_binding)


# === View final table ===
print(intercepts_final)

cat("TB (only first tube spiked) behaves almost identically to TA (non-spiked) → good parallelism, no signal distortion.

TC (every tube spiked) lowers the intercept but not significantly.

TD (1:1 spike + dilution) significantly suppresses signal, indicating likely dilution or matrix effects.")

print(intercepts_binding)

cat("TB again tracks closely with TA, confirming consistent assay response.

TC and TD show increased binding, which is unusual — spiking may be blocking binding sites or altering matrix effects.

The higher binding % may explain the lower final cortisol values, especially in TD")

# === Plot: Intercepts with CI ===
ggplot(intercepts_all, aes(x = Category, y = Estimate, color = Variable)) +
  geom_point(position = position_dodge(width = 0.5), size = 4) +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), 
                width = 0.2, position = position_dodge(width = 0.5)) +
  labs(
    title = "Intercept Estimates for TA–TD",
    subtitle = "With 95% Confidence Intervals",
    y = "Intercept Value",
    x = "Category"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

```




```{r test4 pairs}
# Step 1: Create sample pairs
# e.g., base = TA, spiked = TB with similar Sample ID patterns

# pairing by shared SampleID 
df <- data4 

# Step 1: Extract TA and TB and sample ID
df_pair <- df %>%
  filter(!is.na(Ave_Conc_pg.ml))  %>%
  filter(Category %in% c("A", "D")) %>%
  mutate(SampleID = str_extract(Sample, "\\d+$"))

# Step 2: Pivot to wide format
df_wide <- df_pair %>%
  dplyr::select(SampleID, Category, Ave_Conc_pg.ml, Spike_contribution) 
df_wide <- df_wide %>%
  pivot_wider(names_from = Category, values_from = c(Ave_Conc_pg.ml, Spike_contribution))


# Step 4: Calculate recovery
head(df_wide)
temp <- df_wide %>%
  mutate(Recovery_pct = ((Ave_Conc_pg.ml_D - Ave_Conc_pg.ml_A) / Spike_contribution_D) * 100)

# Step 5: Plot
ggplot(temp, aes(x = SampleID, y = Recovery_pct)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = c(80, 120), linetype = "dashed", color = "darkred") +
  labs(
    title = "Spike Recovery at ELISA Level (pg/mL)",
    x = "Sample ID (TA/TD pair)",
    y = "Recovery (%)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

- reduced detection due to interference or over-saturation. If spike volume is large (e.g. 110 µL into 110 µL sample), it can dilute the original signal below the unspiked concentration

- “Recovery for some high-spike samples was negative, indicating likely signal suppression from matrix effects or over-spiking.”

Can use parallelism to back up conclusion:

- If slope is parallel but intercept drops, it's further evidence of baseline suppression

- Several spiked samples exhibited negative recovery percentages, indicating that measured cortisol concentrations were lower than their unspiked counterparts. This likely reflects signal suppression due to matrix effects or over-dilution from high spike volumes, particularly in the TD group, where spike-to-sample ratios were highest. These results highlight the importance of accounting for spike-related suppression in final quantification strategies.


# Recovery Test4

```{r recov}

# Simulate pairing by shared SampleID (you may need to adjust this step)
df <- data4 

# Step 1: Extract TA and TC and sample ID
df_pair <- df %>%
  filter(!is.na(Ave_Conc_pg.ml))  %>%
  filter(Category %in% c("A", "C")) %>%
  mutate(SampleID = str_extract(Sample, "\\d+$"))

# Step 2: Pivot to wide format
df_wide <- df_pair %>%
  dplyr::select(SampleID, Category, Ave_Conc_pg.ml, Spike_contribution) 
df_wide <- df_wide %>%
  pivot_wider(names_from = Category, values_from = c(Ave_Conc_pg.ml, Spike_contribution))

print(df_wide)
# Step 4: Calculate recovery
temp <- df_wide %>%
  mutate(Recovery_pct = ((Ave_Conc_pg.ml_C - Ave_Conc_pg.ml_A) / Spike_contribution_C) * 100)

# Step 5: Plot
ggplot(temp, aes(x = SampleID, y = Recovery_pct)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = c(80, 120), linetype = "dashed", color = "darkred") +
  labs(
    title = "Spike Recovery at ELISA Level (pg/mL)",
    x = "Sample ID (TA/TC pair)",
    y = "Recovery (%)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r recov test4}


# Simulate pairing by shared SampleID (may need to adjust this step)
df <- data4 
head(data4)

# Step 1: Extract TA and TB and sample ID
df_pair <- df %>%
  filter(!is.na(Ave_Conc_pg.ml))  %>%
  filter(Category %in% c("A", "B")) %>%
  mutate(SampleID = str_extract(Sample, "\\d+$"))

# Step 2: Pivot to wide format
df_wide <- df_pair %>%
  dplyr::select(SampleID, Category, Ave_Conc_pg.ml, Spike_contribution) 
df_wide <- df_wide %>%
  pivot_wider(names_from = Category, values_from = c(Ave_Conc_pg.ml, Spike_contribution))


# Step 4: Calculate recovery
temp <- df_wide %>%
  mutate(Recovery_pct = ((Ave_Conc_pg.ml_B - Ave_Conc_pg.ml_A) / Spike_contribution_B) * 100)


# Step 5: Plot
ggplot(temp, aes(x = SampleID, y = Recovery_pct)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = c(80, 120), linetype = "dashed", color = "darkred") +
  labs(
    title = "Spike Recovery at ELISA Level (pg/mL)",
    x = "Sample ID (TA/TB pair)",
    y = "Recovery (%)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```





