---
title: "Analysis - raw vals test4"
author: "Paloma"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---
# Introduction

```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries, echo = FALSE, warning = FALSE, message=FALSE}
# Install libraries

library(knitr)
library(RColorBrewer)
library(stats)
library(coefplot)
library(arm)
library(bbmle)
library(plyr)
library(dplyr)
library(ggpmisc)
library(ggplot2)
```

```{r loading}
# loading info
# Loading data
data <- read.csv("./data/Test4/Data_QC_flagged.csv")
std <- read.csv("./data/Test4/standard_data_test4.csv")
cort_valsA<- read.csv("./data/Test4/Data_cort_values_methodA.csv")
cort_valsB<- read.csv("./data/Test4/Data_cort_values_methodB.csv")
# remove NSB
data <- data[!(data$Sample == "NSB"),]

#Inclusion of standard readings in plot
data1 <- data[,c("Sample","Wells", "Category", "Binding.Perc", "Ave_Conc_pg.ml", "Weight_mg", "Buffer_nl", "Spike", "Conc_pg.ml", "CV.Perc", "Dilution_sample")]
std1 <- std[,c("Sample", "Wells", "Category", "Binding.Perc", "Backfit")]
#std1$Binding.Perc <- std1$Binding.Perc*100
std1$Spike <- NA
std1$Buffer_nl <- NA
std1$Category <- "Standard"
std1$Weight_mg <- NA
colnames(std1)[5]<- c("Ave_Conc_pg.ml")

data_std <- join(data1, std1, type = "full") 

```

# Summary of results (all experiments)

```{r expABCD.binding.perc, fig.width=7, fig.height=6}
# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data_std, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Binding.Perc)) 
a$Sample_num <- 1:length(a$Sample)

ggplot(a, aes(x = log(Ave_Conc_pg.ml), 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = TRUE, linewidth = 0.3, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.2, linetype = 2, alpha = 0.2, 
              aes(group = Category)) +
  geom_point(size = 2) +
  facet_wrap(vars(Category)) +
  theme_minimal() + 
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
 labs(
  #title = "All serial dilutions and standard: observed binding %",
#       subtitle = "Complete dataset",
       x = "Log(Concentration_pg.ml)", y = "Binding %") +
  scale_y_continuous(n.breaks = 10) 

 
ggplot(a, aes(x = (Ave_Conc_pg.ml), 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = TRUE, linewidth = 0.3, alpha = 0.3) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.2, linetype = 2, alpha = 0.2, 
              aes(group = Category)) +
  geom_point(size = 2) +
  facet_wrap(vars(Category)) +
  theme_minimal() + 
  geom_text(aes(label = Sample), size = 2, vjust = -1, hjust = 0.5, color = "gray30", alpha = 0.5) +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
 labs(
  #title = "All serial dilutions and standard: observed binding %",
#       subtitle = "Complete dataset",
       x = "Concentration_pg.ml", y = "Binding %") +
  scale_y_continuous(n.breaks = 10) 




ggplot(a, aes(x = log(Ave_Conc_pg.ml), 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, linewidth = 0.3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.2, linetype = 2, alpha = 0.2, 
              aes(group = Category)) +
  geom_point(size = 2) +
  theme_minimal() + 
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 0.8) +  
 labs(
  #title = "All serial dilutions and standard: observed binding %",
#       subtitle = "Complete dataset",
       x = "Log(Concentration_pg.ml)", y = "Binding %") +
  scale_y_continuous(n.breaks = 10) 







exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1"))
exp$Sample_num <- 1:length(exp$Sample)

ggplot(exp, aes(x = Sample_num, 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = TRUE, linewidth = 0.2, alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.2, linetype = 2, alpha = 0.2, aes(group = exp$Category)) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  # Add horizontal line 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 3.4, label.y = "bottom", label.x = "right", vjust=-7, hjust = 0.8) +
 # geom_abline(slope = 14.72, intercept = -3.6, linetype = "dashed", color = "lightblue") + # ideal line
  labs(title = "All serial dilutions and standard: observed binding %",
       subtitle = "removed high CV samples, lm = 3 curve",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) + 
    coord_cartesian(ylim = c(0, 105)) +
  theme_minimal() 

```

```{r}

a <- filter(data_std, !is.na(Binding.Perc)) 
a <- a %>%
  filter(!Sample %in% c("B0", "BE", "TA7"))
a$Sample_num <- 1:length(a$Sample)


ggplot(a, aes(x = Binding.Perc, 
                 y = CV.Perc, color = Category)) +
 # geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, linewidth = 0.5, alpha = 0.3) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_vline(xintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_vline(xintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 10, linetype = "dashed", 
             color = "firebrick", linewidth = 1) +  
  labs(title = "Binding and CV limits",
       subtitle = "Complete dataset",
       x = "Binding %", y = "CV %") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```


```{r expABCD.conc.expect.vs.observed}
a <- filter(data_std, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Binding.Perc)) 
a$Sample_num <- 1:length(a$Sample)

a$Exp <- c(4617, 2308.5, 1154.25, 577.125, 288.5625, 144.2813, 72.1406,
           5134, 2567, 1283.5, 641.75, 320.875, 160.4375, 80.2188,
        2467, 1233.5, 616.75, 308.375, 154.1875, 77.0938, 38.5469, 
        4168, 2084, 1042, 521, 260.5, 130.25, 65.125,
        NA, NA, NA, NA, NA, NA, NA)

ggplot(a, aes(x = Exp, 
                 y = (Ave_Conc_pg.ml), color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.06) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray50") +
  stat_poly_eq(aes(x = Exp, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "All serial dilutions: expected vs observed cortisol concentrations",
       subtitle = "Complete dataset",
       x = "Expected vals (pg/mL)", y = "Observed vals (pg/mL)") +
  theme_bw() 


exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1"))


ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml), color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.4, alpha = 0.04) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2, vjust = -1, hjust = 0.5, color = "gray50") +
  stat_poly_eq(aes(x = Exp, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "All serial dilutions: expected vs observed cortisol concentrations",
       subtitle = "Removed low quality samples",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 
``` 

## Experiment A: normal sample dilution

```{r expA.binding.perc}
# Scatter plot of Experiment A, observed binding % 
a <- filter(data_std, Category == "A" | Category =="Standard", !is.na(Binding.Perc)) 
a$Sample_num <- 1:length(a$Sample)

ggplot(a, aes(x = Sample_num, 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  # Add horizontal line 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
  geom_abline(slope = 14.72, intercept = -3.6, linetype = "dashed", color = "lightblue") + # ideal line
  labs(title = "Experiment A: observed binding % of standard vs serial dilution",
       subtitle = "50 mg, 250 uL, no Spike, dilution 1:2",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

```{r expA.conc.expect.vs.observed}

a$Exp <- c(4617, 2308.5, 1154.25, 577.125, 288.5625, 144.2813, 72.1406, NA, NA, NA, NA, NA, NA, NA)

ggplot(a, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment A: expected vs observed cortisol concentrations",
       subtitle = "A=50 mg, 250 uL, no Spike, dilution 1:2",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 

``` 

## Experiment B: spiked sample is serially diluted

```{r expB.binding.perc}

# Scatter plot of Experiment B, observed binding % 
exp <- filter(data_std, Category == "B" | Category =="Standard", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(a$Sample)

ggplot(exp, aes(x = Sample_num, 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  # Add horizontal line 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
  geom_abline(slope = 14.72, intercept = -3.6, linetype = "dashed", color = "lightblue") + # ideal line
  labs(title = "Experiment B: observed binding % of standard vs serial dilution",
       subtitle = "Serial dilution using spiked sample (25uL), 50mg, 250uL, dilution 1:2",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

```{r expB.conc.expect.vs.observed}


# EXPECTED VALS EXPERIMENT B
exp$Exp <- c(5134, 2567, 1283.5, 641.75, 320.875, 160.4375, 80.2188, NA, NA, NA, NA, NA, NA, NA)

ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment B: expected vs observed cortisol concentrations",
       subtitle = "Serial dilution using spiked sample (25uL), 50mg, 250uL, dilution 1:2",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 

### remove TB7
exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1"))

ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment B: expected vs observed cortisol concentrations (removed TB7)",
       subtitle = "Serial dilution using spiked sample (25uL), 50mg, 250uL, dilution 1:2",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 

``` 

## Experiment C: normal sample dilution, and then each sample is spiked

**TC5, 6 and 7 were not spiked (run out of spike)**

```{r expC.binding.perc}

# Scatter plot of Experiment C, observed binding % 
exp <- filter(data_std, Category == "C" | Category =="Standard", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(a$Sample)

ggplot(exp, aes(x = Sample_num, 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  # Add horizontal line 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
  geom_abline(slope = 14.72, intercept = -3.6, linetype = "dashed", color = "lightblue") + # ideal line
  labs(title = "Experiment C: observed binding % of standard vs serial dilution",
       subtitle = "Serial dilution (1:2) using normal sample, then spiked (25uL) afterwards, 50mg, 250uL",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

```{r expC.conc.expect.vs.observed}


# EXPECTED VALS EXPERIMENT C
exp$Exp <- c(2467, 1233.5, 616.75, 308.375, 154.1875, 77.0938, 38.5469, NA, NA, NA, NA, NA, NA, NA)

ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment C: expected vs observed cortisol concentrations",
       subtitle = "Serial dilution (1:2) using normal sample, then spiked (25uL) afterwards, 50mg, 250uL",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 



exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1"))

ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment C: expected vs observed cortisol concentrations",
       subtitle = "Removed low quality samples. Serial dilution (1:2) using normal sample, then spiked (25uL) afterwards, 50mg, 250uL",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 

``` 

## Experiment D: 50% sample, 50% spike, and dilution

```{r expD.binding.perc}

# Scatter plot of Experiment D, observed binding % 
exp <- filter(data_std, Category == "D" | Category =="Standard", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(a$Sample)

ggplot(exp, aes(x = Sample_num, 
                 y = Binding.Perc, color = Category)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 80, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  # Add horizontal line 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
  geom_abline(slope = 14.72, intercept = -3.6, linetype = "dashed", color = "lightblue") + # ideal line
  labs(title = "Experiment D: observed binding % of standard vs serial dilution",
       subtitle = "Serial dilution using spiked sample (110uL), 20mg, 110uL, dilution 1:2",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

```{r expD.conc.expect.vs.observed}
# EXPECTED VALS EXPERIMENT D
exp$Exp <- c(4168, 2084, 1042, 521, 260.5, 130.25, 65.125, NA, NA, NA, NA, NA, NA, NA)

ggplot(exp, aes(x = (Exp), 
                 y = Ave_Conc_pg.ml)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment D: expected vs observed cortisol concentrations",
       subtitle = "Serial dilution using spiked sample (110uL), 20mg, 110uL, dilution 1:2",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 


exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1"))

ggplot(exp, aes(x = (Exp), 
                 y = (Ave_Conc_pg.ml))) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, alpha = 0.05, color = "salmon") +
  geom_point(size = 2, color = "salmon") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = (Exp), label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Experiment D: expected vs observed cortisol concentrations",
       subtitle = "Serial dilution using spiked sample (110uL), 20mg, 110uL, dilution 1:2",
       x = "Expected vals", y = "Observed vals") +
  theme_bw() 

``` 

## Experiment E: testing precision

different weights: 6, 9 and 12 mg, 3 samples for each weight, 2 duplicates for each sample

```{r expE-bindingPerc}
# Scatter plot of Experiment D, observed binding % 


exp <- filter(data_std, Category == "P") %>%
  filter(!Sample %in% c("TP3A")) 


exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  summarise(
    CV_BP = round((sd(Binding.Perc) / mean(Binding.Perc)) * 100, 1), 
    mean_BP = mean(Binding.Perc, na.rm = TRUE),
    sd_BP = sd(Binding.Perc, na.rm = TRUE)
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = mean(Binding.Perc, na.rm = TRUE),
    SD_BP = sd(Binding.Perc, na.rm = TRUE),
    CV_BP = (SD_BP / Mean_BP) * 100
  )
print (all)


ggplot(exp, aes(x = Sample, 
                 y = Binding.Perc, color = factor(Weight_mg))) +
   geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +   geom_smooth(method = "lm", se = TRUE, linewidth = 0.8, alpha = 1, color = "lightblue") +
  geom_point(size = 3, alpha = 1) +
  labs(title = "Experiment E: observed binding % by sample",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(ylim = c(0, 80)) +
  theme_minimal() +
  annotate("text", x = 2, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 5, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 7.5, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 2, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 5, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 7.5, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 


exp <- filter(data_std, Category == "P", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  summarise(
    CV_BP = round((sd(Binding.Perc) / mean(Binding.Perc)) * 100, 1), 
    mean_BP = mean(Binding.Perc, na.rm = TRUE),
    sd_BP = sd(Binding.Perc, na.rm = TRUE)
  )

ggplot(exp, aes(x = Weight_mg, 
                 y = Binding.Perc)) +
   geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
  geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  labs(title = "Experiment E: observed binding % by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  coord_cartesian(xlim = c(4,14), ylim = c(0, 80)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)


exp <- filter(data_std, Category == "P",  Sample != "TP3A", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  summarise(
    CV_BP = round((sd(Binding.Perc) / mean(Binding.Perc)) * 100, 1), 
    mean_BP = mean(Binding.Perc, na.rm = TRUE),
    sd_BP = sd(Binding.Perc, na.rm = TRUE)
  )

ggplot(exp, aes(x = Weight_mg, 
                 y = Binding.Perc)) +
   geom_hline(yintercept = 20, linetype = "dashed", 
             color = "gray", linewidth = 1) +  
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  #geom_hline(yintercept = 80, linetype = "dashed", 
   #          color = "gray", linewidth = 1) +  
 
  labs(title = "Experiment E: observed binding % by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample, removed low quality samples",
       x = "Sample weight (mg)", y = "Binding Percentage") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14), ylim = c(0, 80)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)




```

### Using raw Conc_pg.ml vals estimated by MyAssays
I.e. not controlling for weight or dilution

**TP3A was diluted using 220uL buffer, while all others used 60 uL **

```{r expE-CortVals, fig.width=8, fig.height=6}
# Scatter plot of Experiment D, observed binding % 
exp <- filter(data_std, Category == "P") 
exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  summarise(
    CV_BP = round((sd(Conc_pg.ml) / mean(Conc_pg.ml)) * 100, 1), 
    mean_BP = mean(Conc_pg.ml, na.rm = TRUE),
    sd_BP = sd(Conc_pg.ml, na.rm = TRUE)
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = mean(Conc_pg.ml, na.rm = TRUE),
    SD_BP = sd(Conc_pg.ml, na.rm = TRUE),
    CV_BP = (SD_BP / Mean_BP) * 100
  )


ggplot(exp, aes(x = Weight_mg, 
                 y = Conc_pg.ml, color = Sample)) +
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  labs(title = "Experiment E: concentration vals % by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = "Conc_pg.ml") +
  scale_y_continuous(n.breaks = 10) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)
```

**TP1C has a lower value = I lost sample during extraction, probably tube was broken or cap was not tight enough **

Below I removed a duplicate for sample TP1B as it is a clear outlier, and TP3A as it has a different dilution

```{r expE-CortVals-filtered, fig.width=8, fig.height=6}
# Scatter plot of Experiment D, observed binding % 
exp <- filter(data_std, Category == "P", Wells != "F10", Sample != "TP3A") 
exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  summarise(
    CV_BP = round((sd(Conc_pg.ml) / mean(Conc_pg.ml)) * 100, 1), 
    mean_BP = mean(Conc_pg.ml, na.rm = TRUE),
    sd_BP = sd(Conc_pg.ml, na.rm = TRUE)
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = mean(Conc_pg.ml, na.rm = TRUE),
    SD_BP = sd(Conc_pg.ml, na.rm = TRUE),
    CV_BP = (SD_BP / Mean_BP) * 100
  )


ggplot(exp, aes(x = Weight_mg, 
                 y = Conc_pg.ml, color = Sample)) +
 geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  labs(title = "Experiment E: concentration vals % by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample, removed low quality samples",
       x = "Sample weight (mg)", y = "Conc_pg.ml") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 600, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 400, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)
```




