---
title: "Final vals analysis, test3"
output: html_document
date: "2025-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(knitr)
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
library(tidyverse)
library(readr)

```

# Test3

**Plate description**

Combined variables to evaluate effect of weight, dilution, and addition of spike

```{r test3, echo = TRUE}
#path:
data_path  = "./data/Test3"

# Load the dataset
data3 <- read.csv(file.path(data_path,"Data_QC_flagged.csv"))
data <- data3
data$Failed_samples[is.na(data$Failed_samples)] <- "OK"
# Conversion and Constants
data$Buffer_ml <- data$Buffer_nl / 1000
data$TotalVol_well_ml <- data$TotalVol_well_ul / 1000
data$SpikeVol_ml <- data$SpikeVol_ul / 1000
data$Ave_Conc_ug.dl <- data$Ave_Conc_pg.ml / 10000
# Extraction ratio
extraction <- 1 / 0.75

# Spike concentration (pg/mL)
std_conc <- 3139.5

# Method A: No spike correction
data$Final_pg.mg_A <- ((data$Ave_Conc_pg.ml / data$Weight_mg) *
                         extraction * data$Buffer_ml * data$Dilution)

# Method B: Subtract spike reading (std) and apply dilution factor (×2)
data$Final_pg.mg_B <- ifelse(
  data$Spike == 1,
  ((data$Ave_Conc_pg.ml - std_conc) / data$Weight_mg) *
    extraction * data$Buffer_ml * 2,
  # fallback to method A if not spiked
 data$Final_pg.mg_A
)

# Method C: Spike contribution subtracted
data$Spike_contribution <- ((data$SpikeVol_ml * std_conc) / data$TotalVol_well_ml) / data$Dilution

summary(data$Spike_contribution)

data$Final_pg.mg_C <- (
  (data$Ave_Conc_pg.ml - data$Spike_contribution) / data$Weight_mg) *
    extraction * data$Buffer_ml * data$Dilution


# Save result
write.csv(data, file.path(data_path,"Data_cort_values_ABC.csv"), row.names = FALSE)

# View summary
summary(data$Final_pg.mg_A)
summary(data$Final_pg.mg_B)
summary(data$Final_pg.mg_C)

data3 <- data
```


```{r}

# Reshape data for plotting
data_long <- data %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_B, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

ggplot(data_long, aes(x = Sample, y = Final_pg.mg, color = Method, group = Method)) +
  geom_line(aes(group = Sample), color = "gray75", size = 0.5) +
  geom_point(size = 3, alpha = 0.8, position = position_dodge(width = 0.5)) + 
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue",
               "Final_pg.mg_B" = "orange",
               "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A (Traditional)",
               "Method B (Subtraction + x2)",
               "Method C (Spike contribution)")
  ) +
  labs(
    title = "Hair Cortisol by Sample and Calculation Method",
    x = "Sample ID",
    y = "Final Cortisol (pg/mg)",
    color = "Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "top"
  )


```


```{r}
# Reshape to long format
data_long <- data %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

# Plot

# Plot with lines connecting A and D values for each sample
ggplot(data_long, aes(x = Sample, y = Final_pg.mg, color = Method)) +
  geom_line(aes(group = Sample), color = "gray75", size = 0.5) +  # light line connecting methods
  geom_point(size = 3, alpha = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue", 
               "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A", "Method C")
  ) +
  labs(
    title = "Hair Cortisol Concentration by Sample",
    x = "Sample ID",
    y = "Final Cortisol (pg/mg)",
    color = "Calculation Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "top"
  )
```


```{r}
# Filter to include only good quality samples
data_ok <- data %>%
  filter(Failed_samples == "OK")

# Reshape data for plotting
data_long_ok <- data_ok %>%
  pivot_longer(
    cols = c(Final_pg.mg_A, Final_pg.mg_C),
    names_to = "Method",
    values_to = "Final_pg.mg"
  )

print(data_long_ok, n=5)

# Plot for OK samples only
ggplot(data_long_ok, aes(x = Sample, y = Final_pg.mg, color = Method)) +
  geom_line(aes(group = Sample), color = "gray75", size = 0.5) +
  geom_point(size = 3, alpha = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(
    values = c("Final_pg.mg_A" = "steelblue",
             #  "Final_pg.mg_B" = "orange",
               "Final_pg.mg_C" = "darkgreen"),
    labels = c("Method A (Traditional)",
              # "Method B (Subtraction + x2)",
               "Method C (Spike contribution)")
  ) +
  labs(
    title = "Hair Cortisol (Good Samples Only) — Test3",
    x = "Sample ID",
    y = "Final Cortisol (pg/mg)",
    color = "Method"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "top"
  )

```




# Recovery test 3

```{r recov3}
library(dplyr)
# Load the datax
df <- read.csv(file.path(data_path, "Data_Cortisol_Processed.csv"))
data3_processed <- df
# Clean: extract numeric SampleID suffix to help match pairs
df <- df %>%
  mutate(SampleID = str_extract(Sample_comparable, "\\d+$"))

# Step 1: Wide format - match NoSpike and YesSpike using SampleID
df_wide <- df %>%
  filter(Category %in% c("NoSpike", "YesSpike")) %>%
  dplyr::select(SampleID, Category, Ave_Conc_pg.ml)

# Step 3: Identify IDs that have BOTH NoSpike and YesSpike
df_avg <- df_wide %>%
  dplyr::group_by(SampleID, Category) %>%
  dplyr::summarise(Ave_pgml = mean(Ave_Conc_pg.ml, na.rm = TRUE), .groups = "drop")

# Step 3: Identify IDs that have BOTH NoSpike and YesSpike
paired_ids <- df_avg %>%
  dplyr::count(SampleID) %>%
  filter(n == 2) %>%
  pull(SampleID)


df_paired <- df_avg %>%
  filter(SampleID %in% paired_ids)

df_recovery <- df_paired %>%
  pivot_wider(names_from = Category, values_from = Ave_pgml)

# Step : Define known spike contribution (pg/mL)
# Assume 25 µL of 3200 pg/mL added into 50 µL total = (25/1000 * 3200) / (50/1000)
known_spike_pgml <- 1569.75  # = 1600 pg/mL

df_recovery <- df_recovery %>%
  mutate(Recovery_pct = ((YesSpike - NoSpike) / known_spike_pgml) * 100)

df_recovery <- df_recovery %>%
  mutate(Recovery_Class = case_when(
      Recovery_pct < 80 ~ "Under-recovery",
      Recovery_pct > 120 ~ "Over-recovery",
      TRUE ~ "Acceptable"))

# Step 5: Plot
ggplot(df_recovery, aes(x = SampleID, y = Recovery_pct, fill = Recovery_Class)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = c(80, 120), linetype = "dashed", color = "darkred") +
  scale_fill_manual(values = c(
    "Under-recovery" = "red",
    "Acceptable" = "steelblue",
    "Over-recovery" = "orange2"
  )) +
  labs(
    title = "Spike Recovery (%) at ELISA Level",
    subtitle = "Dashed lines show 80–120% acceptable range",
    x = "Sample ID",
    y = "Recovery (%)",
    fill = "Recovery Classification"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Step 4: View or export summary table
print(df_recovery)

```
All samples show recovery ≥ 105%, with most between 105% and 132%

Sample 5, 8, 9, and 11–12 exceed the standard acceptable range of 80–120%

This suggests slightly inflated recovery, potentially due to:

Minor pipetting differences

Matrix effects (e.g., enhanced absorbance in spiked matrix)

Slight overestimation in YesSpike values relative to expected spike contribution


Spike recovery ranged from 105% to 155% across 10 matched sample pairs. Most samples fell near or slightly above the commonly accepted 80–120% range, with the highest recovery observed in Sample 5 (155%). These results indicate that the ELISA assay successfully detects added cortisol with high sensitivity, though minor over-recovery may reflect matrix effects or variation in sample absorbance. Overall, recovery was consistent and supports the validity of the spike-in approach at the ELISA level.



# Precision and accuracy, test3

```{r prec and acc}
df <- data3_processed

# Step 1: Clean and filter only relevant categories
df_sub <- df %>%
  filter(Category %in% c("NoSpike", "YesSpike")) %>%
  mutate(Sample_comparable = as.character(Sample_comparable))

# Step 2: Keep only Sample_comparable IDs with both YesSpike and NoSpike
valid_pairs <- df_sub %>%
  distinct(Sample_comparable, Category) %>%
  dplyr::count(Sample_comparable) %>%
  filter(n == 2) %>%
  pull(Sample_comparable)

df_matched <- df_sub %>%
  filter(Sample_comparable %in% valid_pairs)

# Step 3: Pivot to wide format for recovery calculation
df_wide <- df_matched %>%
  dplyr::select(Sample_comparable, Category, Ave_Conc_pg.ml) 

df_avg <- df_wide %>%
  dplyr::group_by(Sample_comparable, Category) %>%
  dplyr::summarise(Ave_pgml = mean(Ave_Conc_pg.ml, na.rm = TRUE), .groups = "drop")

Spike_contribution <- 1569.75

df_recovery <- df_avg %>%
  pivot_wider(names_from = Category, values_from = Ave_pgml)


# Step 4: Calculate Recovery % and accuracy flag
df_wide <- df_recovery %>%
  mutate(
    Recovery_pct = ((YesSpike - NoSpike) / Spike_contribution) * 100,
    Accuracy_Flag = case_when(
      is.na(Recovery_pct) ~ NA_character_,
      Recovery_pct < 80 ~ "Under-recovery",
      Recovery_pct > 120 ~ "Over-recovery",
      TRUE ~ "Acceptable"
    )
  )

# Step 5: Extract precision values
df_precision <- df %>%
  dplyr::select(Sample, Sample_comparable, CV.Perc, Weight_mg, Buffer_ml) %>%
  mutate(Precision_Flag = case_when(
    is.na(CV.Perc) ~ NA_character_,
    CV.Perc <= 15 ~ "Acceptable",
    TRUE ~ "High CV"
  ))

df_precision$Sample_comparable<- as.character(df_precision$Sample_comparable)
# Step 6: Merge precision with accuracy summary
df_summary <- left_join(df_wide, df_precision, by = "Sample_comparable") %>%
  distinct(Sample_comparable, .keep_all = TRUE) %>%
  dplyr::select(Sample, Weight_mg, Buffer_ml, CV.Perc, Precision_Flag, Recovery_pct, Accuracy_Flag)

df_sorted <- df_summary %>% arrange(Weight_mg, Buffer_ml)
df_sorted$Sample <- 1:nrow(df_sorted)
# Step 7: View or export
print(df_sorted)


# Optional: Save to file
write.csv(df_sorted, file.path(data_path, "precision_accuracy_summary.csv"), row.names = FALSE)

ggplot(df_sorted, aes(x = Weight_mg, y = Recovery_pct,
                      color = Accuracy_Flag, shape = Precision_Flag)) +
  geom_point(size = 4) +
  geom_hline(yintercept = c(80, 120), linetype = "dashed", color = "darkgray") +
  labs(
    title = "Spike Recovery (%) vs Hair Weight (mg)",
    subtitle = "Dashed lines show 80–120% acceptable range",
    x = "Hair Weight (mg)",
    y = "Recovery (%)",
    color = "Accuracy",
    shape = "Precision"
  ) +
  theme_minimal()           
           


```

