---
title: "Cortisol Concentration Values, Test3"
author: "Paloma Contreras"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

# Summary

**Final cortisol value calculations** were conducted using three methods:

- **Standard Method (Method A)**: Calculates cortisol concentration without correction for spiked samples.

- **Spike-Corrected Method (Method B)**: Adjusts for spiked samples to account for addition of a known amount of cortisol, following Nist et al. 2020.

- **Sam's Method (Method C)**: Adjusts for spiked samples using a different equation

**Results**: As we see below, the formula used by Nist et al. results in negative values, which would mean that there is no cortisol in original samples. This could be an artifact of an extremely high absorbance level caused by an excessive amount of spike. Non-spiked samples, on the other hand, result in values that are within the range found in similar studies of cortisol in human hair. 

<table>
Summary                 Nist et al.     (A) Standard      (B) Spike-Corrected        (C) Sam's     
---------------         ------------   ---------------    ---------------------      -----------------
Mean cort conc (pg/mg)      ---         16.796              -0.1866                    9.568
Median cort conc (pg/mg)    ---           10.801            4.4386                     10.058
Range cort conc (pg/mg)     ---           2.7 to 60.4       -30.7005 to 12.0647         2.786 to 22.566
</table>


<table>

Weight (mg) of my samples
---------------           -------------  
Range                     11 to 37.1
Mean                      23.54
Median                    22.4                 
                                                             
</table>

**Conclusions**: 
After accounting for differences in dilution and weight, our results suggest future Assays should use the optimal parameters listed below: 

- Dilution of 250uL is preferable over 60uL 
- Non-spiked samples seem to generate expected results


**Concerns**

- Spike results in unrealistic values
- Could be explained by the higher weight of our samples 
- Dilution of 250uL results in values that are twice as big as with 60uL, but they should be very similar or at least overlap

## Explanation of each variable used in calculations

- Ave_Conc_pg/ml: average ELISA reading per sample in pg/mL

- Weight_mg: hair weight in mg

- Buffer_nl: assay buffer volume in nL → we convert to mL

- Spike: binary indicator (1 = spiked sample)

- SpikeVol_uL: volume of spike added in µL

- Dilution: dilution factor (already accounted for)

- Vol_in_well.tube_uL: total volume in well/tube in µL (for spike correction)

- std: standard reading value

- extraction: methanol volume ratio = vol added / vol recovered (e.g. 1/0.75 ml)


# Cortisol concentration calculations
Input is data with low quality samples flagged, but they get removed before continuing with calculations. 
```{r define dataset, echo =FALSE}
# DATA SET 
current_test <- "Test3"
data_path <- file.path("./data", current_test)
df <- read.csv(file.path(data_path,"Data_QC_flagged.csv"))
```

```{r libraries, echo = FALSE, message=FALSE}
library(knitr)
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
```
Parameters and unit transformations: 
```{r parameters}
# Define volume of methanol used for cortisol extraction
# vol added / vol recovered (mL)
extraction <- 1 / 0.75

# Reading of spike standard and conversion to ug/dl
std <- 3139.5 # test 3 backfit
std_ul.dl <- std / 10000 # std in ul/dl

# Creating variables in indicated units

df$Buffer_ml <- c(df$Buffer_nl/1000) # dilution (buffer)
df$Ave_Conc_ug.dl <- c(df$Ave_Conc_pg.ml/10000) # Transform to μg/dl from assay output
```

```{r echo=FALSE}
# remove unnecessary and low quality data 
data <- df %>%
    filter(CV.Perc < 15) %>% 
    filter(Binding.Perc < 80 & Binding.Perc > 20) %>%
    dplyr::select(Wells, Sample, Category, Binding.Perc, Ave_Conc_pg.ml, Ave_Conc_ug.dl, Weight_mg, Buffer_ml, Spike, SpikeVol_ul, Dilution, TotalVol_well_ul, Failed_samples, Sample_comparable) 

# remove duplicates
data.og <- data[!is.na(data$Binding.Perc), ]
kable(head(data.og, 3))
```

## (A) Standard Calculation

Formula: 

((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r A}
##################################
##### Calculate final values #####
##################################

data$Final_conc_pg.mg <- c(
    ((data$Ave_Conc_ug.dl) / data$Weight_mg) *    # A/B *
      extraction *                                # C/D  *     
      data$Buffer_ml * 10000 )                    # E * 10000


```

```{r echo = FALSE}
# summary for all samples
data <- data[order(data$Sample),]
write.csv(data, file.path(data_path, "Data_cort_values_methodA.csv"), row.names = F)
cat("Summary of all samples")
summary(data$Final_conc_pg.mg)
kable(tail(data, 4))
dSpike <- data
```

## (B) Accounting for Spike

We followed the procedure described in **Nist et al. 2020**:

"Thus, after pipetting 25μL of standards and samples into the appropriate wells of the 96-well assay plate, we added 25μL of the 0.333ug/dL standard to all samples,
resulting in a 1:2 dilution of samples. The remainder of the manufacturer’s protocol was
unchanged. We analyzed the assay plate in a Powerwave plate reader (BioTek, Winooski,
VT) at 450nm and subtracted background values from all assay wells. In the calculations, we
subtracted the 0.333ug/dL standard reading from the sample readings. Samples that resulted
in a **negative number were considered nondetectable**. We converted cortisol levels from
ug/dL, as measured by the assay, to pg/mg—based on the mass of hair collected and
analyzed using the following formula:

A/B * C/D * E * 10,000 * 2 = F

where 
- A = μg/dl from assay output; 
- B = weight (in mg) of collected hair; 
- C = vol. (in ml) of methanol added to the powdered hair; 
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down; 
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract; 10,000 accounts for changes in metrics; 2 accounts for the dilution factor after addition of the spike; and 
- F = final value of hair cortisol concentration in pg/mg"


```{r B-calc}
##################################
##### Calculate final values #####
##################################

dSpike$Final_conc_pg.mg <- 
  ifelse(
    dSpike$Spike == 1,                     ## Only spiked samples
      ((dSpike$Ave_Conc_ug.dl - (std_ul.dl)) / # (A-spike) 
        dSpike$Weight_mg)                  # / B
        * extraction *                     # C / D
        dSpike$Buffer_ml * 10000 * 2,      # E * 10000 * 2
        dSpike$Final_conc_pg.mg  
    )
```

```{r B-save, echo = FALSE}
write.csv(dSpike, file.path(data_path, "Data_cort_values_methodB.csv"), row.names = F)

cat("Summary of all samples")
# summary for all samples
summary(dSpike$Final_conc_pg.mg)

cat("Summary without spiked samples")
dSpikeSub <- data[c(data$Spike == 0), ]
summary(dSpikeSub$Final_conc_pg.mg)

kable(tail(dSpike, 3))
```

## (C) Sam's calculation

Developed using Sam's advice and logic. To facilitate the understanding of what is going on, here I do not transform the output values from pg/ml to ug/dL (as done in A and B). 

- Step 1: Calculate contribution of the spike
- Step 2: Substract spike from plate reading values and calculate final values accounting for dilution of the sample, weight, and reconstitution

**Step 1**: Calculate contribution of spike

 X * Y / Z / SPd = SP

- SP = final value of spike contribution in pg/mL
- X = volume of spike added (mL)
- Y = concentration of the spike added (pg/mL)
- SPd = if serially diluted, dilution factor for the spike (i.e: 1, 2, 4, 8, etc.)
- Z = total volume (mL) in the well or tube, if spike is added before loading the plate (sample + spike)

```{r C-calc}
# Transforming units
data$SpikeVol_ml <- data$SpikeVol_ul/1000           # X to mL
data$TotalVol_well_ml <- data$TotalVol_well_ul/1000 # Z to mL
      # SPd = dilution (in this case, is 1 for all)

 # Calculate spike contribution to each sample
      ##  ( Spike vol. x Spike Conc.)
      ##   ------------------------  / dilution = Spike contribution
      ##        Total vol. 

data$Spike.cont_pg.ml <- (((data$SpikeVol_ml * std ) / # X * Y /
                            data$TotalVol_well_ml) /   # Z /
                              data$Dilution)           # SPd 
summary(data$Spike.cont_pg.ml)
```

```{r echo = FALSE}
dSpiked <- data
cat("The reading for standard 1 in this plate is", std)
cat("The total contribution of the Spike to each sample is 1569.75 pg/mL")
```

**Step 2 **: Substract spike and calculate final values

((A - **SP**)/B) * (C/D) * E * **SLd**  = F

- A = pg/ml from assay output;
- *SP* = spike contribution (in pg/ml) 
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r}
##################################
##### Calculate final values #####
##################################
dSpiked$Final_conc_pg.mg <- 
   ((dSpiked$Ave_Conc_pg.ml - dSpiked$Spike.cont_pg.ml) / # (A - spike) 
   dSpiked$Weight_mg) *                                   # / B *
   extraction *                                           # C / D
   dSpiked$Buffer_ml                 # E 
dSpiked[ , c("Sample", "Ave_Conc_pg.ml", "Buffer_ml","Spike.cont_pg.ml", "Final_conc_pg.mg")]
```

```{r echo = FALSE}
write.csv(dSpiked, file.path(data_path, "Data_cort_values_methodC.csv"), row.names = F)

cat("Summary for all samples:")
summary(dSpiked$Final_conc_pg.mg)
kable(tail(dSpiked[!is.na(dSpiked$Final_conc_pg.mg) , c("Sample", "Final_conc_pg.mg", "Ave_Conc_pg.ml", "Spike.cont_pg.ml", "Binding.Perc", "Weight_mg", "Buffer_ml", "SpikeVol_ul", "Dilution", "TotalVol_well_ul")], 10))
```

# Plots 

## (A) Standard Calculation

Final cortisol concentrations not accounting for spike. Tags are sample numbers. 

*Expected results*: a straight horizontal line showing that I obtained same cortisol concentration value in the Y axis, across different sample weights. 

```{r echo = FALSE, warning=FALSE }
# scatterplot method A
data$Spike <- replace(data$Spike, data$Spike == 1, 'Yes')
data$Spike <- replace(data$Spike, data$Spike == 0, 'No')
data$Buffer <- data$Buffer_ml
data$Buffer <- replace(data$Buffer, data$Buffer == 0.06, '60 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.11, '110 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.22, '220 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.25, '250 uL')


ggplot(data, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 3.5, alpha = 0.85) +  
   geom_text(aes(label = Sample), size = 2.5, vjust = -0.65, hjust = -0.18) +
    theme_minimal() +  
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
#  xlim(0,52) +
  labs(
    title = "(A) Standard Calculation",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)") + 
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  ) 
```

## (B) Accounting for Spike

Final cortisol concentrations accounting for Spike as instructed in Nist et al. 2020. 

*Expected results*: lower values than in the previous plot for the spiked samples, but not as low as negative samples (for all of them). Spiked and non-spiked samples should be aligned (same concentration across different weights)

```{r bplot, echo = FALSE, warning=FALSE }

dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 1, 'Yes')
dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 0, 'No')
dSpike$Buffer <- dSpike$Buffer_ml
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.06, '60 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.11, '110 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.22, '220 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.25, '250 uL')

# scatterplot
ggplot(dSpike, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 3.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
#  xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(B) Calculation Accounting for Spike",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )

```

## (C) Sam's calculation

Final cortisol concentration values using new method. 

*Expected results*: one unique horizontal line, regardless of the spiking status and dilution. We see this line for the spiked samples that were reconstituted using 60 uL (i.e, the most concentrated samples). Perhaps the 250uL samples, by being more diluted and having a larger volume, present more variation if the cort distribution within the well/tube is not homogeneous.  
*Note* that samples seem to be less aligned or more separated from each other than in previous plots. This is due to a difference in scale (A has values 0 to 60, while here all values fall between 2.5 and 22.5). 
```{r cplot, echo = FALSE, warning=FALSE }

dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 1, 'Yes')
dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 0, 'No')
dSpiked$Buffer <- dSpiked$Buffer_ml
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.06, '60 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.11, '110 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.110, '110 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.25, '250 uL')

dSpiked$Group <- interaction(dSpiked$Buffer, dSpiked$Spike, sep = "_")

# scatterplot
ggplot(dSpiked, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 fill = Buffer,
                 shape = Spike, 
                 color = Buffer)) +
    geom_smooth(method = "lm", 
              se = TRUE,
              alpha = 0.2,
              linewidth = 0.4,
              color = "gray60"
                ) +
    geom_point(size = 3.5,  
               alpha = 0.85) + 
    geom_text(aes(label = Sample), 
              size = 3, vjust = -1, hjust = -0.1) +
    geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red3") +
  theme_minimal() +  
   # ylim(-26,30) +
  #xlim(0,52) +
  labs(
    title = "(C) Sam's method ",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"
    ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )

```

# Evaluation using method C

```{r all samples, message=FALSE, echo = FALSE}
# All samples 
data2 <- dSpiked

#### fit models ####

# model all
model.all <- lm(Final_conc_pg.mg ~ Weight_mg, 
              data = data2)
r_squared.all <- summary(model.all)$r.squared

# Calculate residuals
residuals.all <- residuals(model.all)

# Mean Absolute Error
mae.all <- mean(abs(residuals.all))      
# Standard Deviation of Residuals
std_dev.all <- sd(residuals.all)

```

```{r plot1, echo = FALSE}
ggplot(data2, aes(y = Final_conc_pg.mg, 
                  x = Weight_mg)) +
  geom_point(size = 2.5, color = "darkorange2") +  
  geom_text(label = c(data2$Sample), nudge_y = 0.75, nudge_x = -0.5) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.3) + 
  geom_hline(yintercept = mean(data2$Final_conc_pg.mg), 
             color = "lightblue3",
             linetype = "dashed") +
  theme_minimal() +  
  xlim(5, max(data2$Weight_mg) + 4) +
  ylim(0, max(data2$Final_conc_pg.mg)+4) + 
  labs(
    title = "Final Cort Concentration vs Weight
    (All samples) method C",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) +
  # Add R^2 annotation
  annotate("text", x = 28, 
           y = max(data2$Final_conc_pg.mg),
           label = paste("R² =", round(r_squared.all, 4)), 
           size = 4, color = "black") + 
    annotate("text", x = 30, 
           y = max(data2$Final_conc_pg.mg) - 1.5,
           label = paste("Mean Residuals =", round(mae.all, 3)), 
           size = 4, color = "black") +
   annotate("text", x = 30, 
           y = max(data2$Final_conc_pg.mg) - 3,
           label = paste("SD Residuals =", round(std_dev.all, 3)), 
           size = 4, color = "black")
```

```{r separate reconstitution vols, echo = FALSE}
# Calc modelo 250 and 60 ul

#two datasets, separated by dilution
data2.06 <- data2[data2$Buffer == "60 uL", ]
data2.25 <- data2[data2$Buffer == "250 uL", ]

#### fit models ####

# model Buffer = 0.06
model06 <- lm(Final_conc_pg.mg ~ Weight_mg, 
              data = data2.06)
r_squared06 <- summary(model06)$r.squared

# model Buffer = 0.25
model25 <- lm(Final_conc_pg.mg ~ Weight_mg, 
              data = data2.25)
r_squared25 <- summary(model25)$r.squared

# Calculate residuals
residuals06 <- residuals(model06)
residuals25 <- residuals(model25)

# Quantify residuals
# Mean Absolute Error
mae06 <- mean(abs(residuals06))          
# Standard Deviation of Residuals
std_dev06 <- sd(residuals06)   

# Mean Absolute Error
mae25 <- mean(abs(residuals25))      
# Standard Deviation of Residuals
std_dev25 <- sd(residuals25)                      

```

```{r plot2, message=FALSE, echo = FALSE}
             
# scatterplot
ggplot(data2, aes(y = Final_conc_pg.mg, 
                  x = Weight_mg,
                  color = Buffer,
                  fill = Buffer)) +
  geom_point(size = 2.5, 
             aes(shape = Spike)) +  
  geom_text(label = c(data2$Sample),size = 3, nudge_y = 0.75, nudge_x = -0.5) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.2,
              linewidth = 0.4) + 
  geom_hline(yintercept = mean(data2$Final_conc_pg.mg), 
             color = "gray80",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.06$Final_conc_pg.mg), 
             color = "turquoise2",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.25$Final_conc_pg.mg), 
             color = "salmon",
             linetype = "dashed") +
  theme_minimal() +  
  xlim(5, max(data2$Weight_mg) + 4) +
  ylim(-1.5, max(data2$Final_conc_pg.mg)+4) + 
  labs(
    title = "Final Cort Concentration and Weight
    (All samples) method C", 
    subtitle = "Grouped by reconstitution volume",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  )  +
  # Add R^2 annotation
     annotate("text", x = min(data2$Weight_mg) - 5.2, 
           y = min(data2$Final_conc_pg.mg) + 1.0,
           label = paste("60 uL"), 
           size = 3.8, color = "turquoise4") +
  annotate("text", x = min(data2$Weight_mg) - 4.45 , 
           y = min(data2$Final_conc_pg.mg - 0.8),
           label = paste("R² =", round(r_squared06, 3)), 
           size = 3.8, color = "turquoise4") +
  annotate("text", x = min(data2$Weight_mg - 1) , 
           y = min(data2$Final_conc_pg.mg) - 2,
           label = paste("Mean Residuals =", round(mae06, 3)), 
           size = 3.8, color = "turquoise4") +
  annotate("text", x = min(data2$Weight_mg) -1.5, 
           y = min(data2$Final_conc_pg.mg) - 3.2,
           label = paste("SD Residuals =", round(std_dev06, 3)), 
           size = 3.8, color = "turquoise4") +
   annotate("text", x = max(data2$Weight_mg) - 5, 
           y = max(data2$Final_conc_pg.mg) + 1.65,
           label = paste("250 uL"), 
           size = 3.8, color = "salmon2") +
  annotate("text", x = max(data2$Weight_mg) - 4, 
           y = max(data2$Final_conc_pg.mg),
           label = paste("R² =", round(r_squared25, 3)), 
           size = 3.8, color = "salmon2") +
  annotate("text", x = max(data2$Weight_mg - 1) , 
           y = max(data2$Final_conc_pg.mg) - 1.2,
           label = paste("Mean Residuals =", round(mae25, 3)), 
           size = 3.8, color = "salmon2") +
  annotate("text", x = max(data2$Weight_mg) -1.5, 
           y = max(data2$Final_conc_pg.mg) - 2.4,
           label = paste("SD Residuals =", round(std_dev25, 3)), 
           size = 3.8, color = "salmon2") 
   
```

The previous figure shows that: 

- results are relatively stable across weights, particularly for the samples where a dilution of 250 uL was used 
- For the 250 uL reconstitutions, about 18% of the variability in Final_conc_pg.mg is explained by Weight_mg 
- For the 60 uL reconstitutionsAbout 6.6% of the variability is explained by Weight_mg. 
- Reconstitution volume affects estimation of cortisol concentration in a significant way: even though final concentration numbers account for differences in the dilutions, the results we observe for each group do not overlap
- the average value when using 250 uL of buffer is twice as big as when using 60 uL


```{r prep-spiked, echo = FALSE, warning = FALSE}
#two datasets, separated by spiked or not
data2.sp <- data2[data2$Spike == "Yes", ]
data2.nsp <- data2[data2$Spike == "No", ]

# model Spike
model.sp <- lm(Final_conc_pg.mg ~ Weight_mg, 
              data = data2.sp)
r_squared.sp <- summary(model.sp)$r.squared
# model non spike
model.nsp <- lm(Final_conc_pg.mg ~ Weight_mg, 
              data = data2.nsp)
r_squared.nsp <- summary(model.nsp)$r.squared
# residuals 
residuals.sp <- residuals(model.sp)
residuals.nsp <- residuals(model.nsp)

# Mean Absolute Error
mae.sp <- mean(abs(residuals.sp))      
# Standard Deviation of Residuals
std_dev.sp <- sd(residuals.sp)

# Mean Absolute Error
mae.nsp <- mean(abs(residuals.nsp))      
# Standard Deviation of Residuals
std_dev.nsp <- sd(residuals.nsp)
```

```{r plot3-spiked vs non-spiked, message=FALSE, echo = FALSE}

# scatterplot

ggplot(data2, aes(y = Final_conc_pg.mg, 
                  x = Weight_mg, 
                  color = Spike, 
                  fill = Spike)) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.2,
              linewidth = 0.5) + 
  geom_hline(yintercept = mean(data2$Final_conc_pg.mg), 
             color = "gray80",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.sp$Final_conc_pg.mg), 
             color = "turquoise4",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.nsp$Final_conc_pg.mg), 
             color = "salmon",
             linetype = "dashed") +
  geom_point(size = 2) +  
  geom_text(label = c(data2$Sample), 
            nudge_y = 0.75, nudge_x = -0.5,
            size = 2.8) +
  theme_minimal() +  
         xlim(5, max(data2$Weight_mg) + 4) +
         ylim(0, max(data2$Final_conc_pg.mg)+4) + 
  labs(
    title = "Final Cort Concentration and Weight, method C",
    subtitle = "Grouped by spiked/non-spiked", 
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) +
  # Add R^2 annotation
  annotate("text", x = max(data2$Weight_mg) - 4.5 , 
           y = min(data2$Final_conc_pg.mg),
           label = paste("R² =", round(r_squared.nsp, 3)), 
           size = 4, color = "turquoise4") +
  annotate("text", x = max(data2$Weight_mg - 1) , 
           y = min(data2$Final_conc_pg.mg) - 1.1,
           label = paste("Mean Residuals =", round(mae.nsp, 3)), 
           size = 4, color = "turquoise4") +
  annotate("text", x = max(data2$Weight_mg) -1.5, 
           y = min(data2$Final_conc_pg.mg) - 2.2,
           label = paste("SD Residuals =", round(std_dev.nsp, 3)), 
           size = 4, color = "turquoise4") +
  annotate("text", x = max(data2$Weight_mg) -4.25, 
           y = max(data2$Final_conc_pg.mg),
           label = paste("R² =", round(r_squared.sp, 3)), 
           size = 4, color = "salmon2") +
    annotate("text", x = max(data2$Weight_mg - 1) , 
           y = max(data2$Final_conc_pg.mg) - 1.1,
           label = paste("Mean Residuals =", round(mae.sp, 3)), 
           size = 4, color = "salmon2") +
  annotate("text", x = max(data2$Weight_mg) -1.5, 
           y = max(data2$Final_conc_pg.mg) - 2.2,
           label = paste("SD Residuals =", round(std_dev.sp, 3)), 
           size = 4, color = "salmon2") 
```



```{r, echo=FALSE, message=FALSE, results = 'hide', include = FALSE}
qqnorm( residuals.all) 
qqline(residuals.all)
plot(density(residuals.all))

qqnorm( residuals25) 
qqline(residuals25)
plot(density(residuals25))

qqnorm( residuals06) 
qqline(residuals06)
plot(density(residuals06))

```

## Optimal dilution (using method C results)

**Error using samples w/0.06 mL buffer**
```{r echo = FALSE}
# Print the results
cat("Mean Absolute Error (MAE) 0.06 mL:", round(mae06, 3), "\n")
cat("Standard Deviation of Residuals 0.06 mL:", round(std_dev06, 3), "\n")
```
**Error using samples w/0.25 mL buffer**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) 0.25 mL:", round(mae25, 3), "\n") 
cat("Standard Deviation of Residuals 0.25 mL:", round(std_dev25, 3), "\n")

## Abs error and std dev. are larger for dilution 0.06 mL
```
From this we conclude that using **a 60 uL dilution produces more accurate/more consistent results**

**Error using spiked samples only**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) ALL:", round(mae.sp, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.sp, 3), "\n")
```
**Error using non-spiked samples only**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) ALL:", round(mae.nsp, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.nsp, 3), "\n")
```
**Error using all samples**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) ALL:", round(mae.all, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.all, 3), "\n")
```
