---
title: "Calculation Final Values, Test7"
author: "Paloma Contreras"
date: "`r Sys.Date()`"
output: workflowr::wflow_html

---

# Summary

**I. Cortisol values (pg/mg)** 


|                        | Min. | 1st Qu. | Median | Mean | 3rd Qu. | Max. | NA's |
|------------------------|------|---------|---------|---------|---------|-------|------|
|**Good quality samples (n=11)**| 10.31 |  43.77  | 60.09 | 55.02 | 64.91| 87.61  | 0|
|**All samples (n=27)**|  10.31 |  55.13  | 69.95 | 303.97 | 100.10 |3055.36   |    1 |


**II. Plate description**

Category	  Description
---------  ---------------------------------------------------------------
P10         Paloma, 10 mg, 220 uL
P20         Paloma, 20 mg, 220 uL
T20         Tina,   20 mg, 220 uL
Pool        Paloma, 20mg,  250 uL
Std-B       Serial dilution of standard 1 (3200 pg/mL)


**III. Results:**  

- Mean Intra-assay CV: 12.02% 

- Mean Intra-assay CV after removing low quality samples: 4.75%

- Inter-assay CV: **-- ** 

(Bindings for 20mg sample diluted in 250 uL, no spike: 64.8% and 48% in test3 and test4, respectively)


**IV. Conclusions**: 



**V. Concerns:**
Tina continues to fall under the curve even after a dilution 32

# Loading and inspecting data

```{r Loading files, echo = FALSE, message=FALSE}
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
library(knitr)
library(here)

#path:
data_path <- "./data/Test7"
cat("Input:")
print(file.path(data_path, "Data_QC_flagged.csv"))  # Check the full path
df <- read.csv(file.path(data_path, "Data_QC_flagged.csv"))
df$Buffer_ml <- df$Buffer_nl / 1000

cat("Identify and flag bad quality samples")

# remove outliers
# label samples by quality

df$Failed_samples[is.na(df$Failed_samples)] <- "OK"
table(df$Failed_samples)
df$Failed_samples[df$Failed_samples == "ABOVE 80% binding"] <- "Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;ABOVE 80% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;UNDER 20% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV"] <- "High CV"
df$Failed_samples[df$Failed_samples == "UNDER 20% binding"] <- "Out of curve"

df$Failed_samples <- factor(
  df$Failed_samples,
  levels = c(
    "OK",
    "Out of curve",
    "High CV",
    "High CV & Out of curve"
  )
)

cat("Summary CV for all samples:")
summary(df$CV.Perc)

cat("Summary CV for good quality samples only:")
temp <- df %>% 
  filter(Failed_samples == 
           "OK")
summary(temp$CV.Perc)


# remove unnecessary information 
data <- df %>%
    dplyr::select(Wells, Sample, Category, Binding.Perc, 
                  Ave_Conc_pg.ml, Weight_mg, 
                  Buffer_ml, Dilution_sample, 
                 Extraction_ratio,  
                  Failed_samples) 


data <- data[!is.na(data$Binding.Perc), ] # remove duplicates

```

# Final value calculation

Calculate final values accounting for dilution of the sample (from myassays.com), weight, and reconstitution

((A)/B) * (C/D) * E  = F

- A = pg/ml from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r}
##################################
##### Calculate final values #####
##################################

data$Final_pg.mg <- 
     (data$Ave_Conc_pg.ml / data$Weight_mg) *      # A / B *
     data$Extraction_ratio *                       # C / D *
     data$Buffer_ml                                # E 
```

## Explanation of each variable used in calculations

- Ave_Conc_pg/ml: average ELISA reading per sample in pg/mL

- Weight_mg: hair weight in mg (here we used 10 and 20 mg)

- Buffer_ml: assay buffer volume in mL (here we used 220 uL)

- Extraction_ratio: methanol volume ratio = vol added / vol recovered (here we used 1.4/1 ml)

Dilution: dilution factor is already included in values obtained from myassays.com

# Results 
Cortisol concentration values are shown in pg/mg

```{r echo = FALSE}
write.csv(data, file.path(data_path, "Data_cort_values.csv"), row.names = F)

# summary for all samples
data <- data %>%
  filter(!Sample %in% c("B0", "BE", "NSB")) 
cat("Summary for all samples:")
summary(data$Final_pg.mg)

cat("Summary for good quality samples only:")
temp <- data %>% 
  filter(Failed_samples == 
           "OK")
summary(temp$Final_pg.mg)

```

## Plot: results grouped by category
```{r plot-by-categ, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(data, aes(y = Final_pg.mg, 
                  x = 1:length(Sample),
                  color = Category,
                  fill= Category)) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.2) + 
  geom_hline(yintercept = 0, 
             color = "lightblue3",
             linetype = "dashed") +
    geom_point(size = 2.5) +  
  ylim(0, max(data$Final_pg.mg)) +
   geom_text(label = c(data$Sample), nudge_y = 100, nudge_x = 0, size = 2) +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration by Category
    (All samples)",
    y = "Final Concentration (pg/mg)",
    x = "Sample number"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) 
```

## Plot: results grouped by category without outliers
```{r plot-by-categ-2, echo = FALSE, warning = FALSE, message = FALSE}

ggplot(data, aes(y = Final_pg.mg, 
                  x = 1:length(Sample),
                  color = Category,
                  fill= Category)) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.2) + 
  geom_hline(yintercept = 0, 
             color = "lightblue3",
             linetype = "dashed") +
    geom_point(size = 2.5) +  
       ylim(0,135) +
  geom_text(label = c(data$Sample), nudge_y = 3, nudge_x = 0, size = 2.2) +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration by Category
    (All samples)",
    y = "Final Concentration (pg/mg)",
    x = "Sample number"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) 
```

## Plot: results grouped by data quality
```{r echo = FALSE, warning=FALSE }
# scatterplot 

data$Buffer <- data$Buffer_ml

data$Buffer <- replace(data$Buffer, data$Buffer == 0.22, '220 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.25, '250 uL')

ggplot(data, aes(y = Final_pg.mg, 
                 x = 1:length(Sample), 
                 color = Failed_samples)) +
  geom_point(size = 2.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 2.3, vjust = -1, hjust = 0.6) +
  theme_minimal() +  
 #    ylim(-26,30) +
  #   xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "Final cort values",
    y = "Final Concentration (pg/mg)",
    x = "Sample" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  )
```


## Error evaluations, grouped by data quality 
```{r models-good-quality, message=FALSE, echo = FALSE}

cat("Models used with respect to weight: 
      lm(Final.pg.mg ~ Weight_mg, 
                  data = data)")

#### fit models ####
# model Spiked
model.all <- lm(Final_pg.mg ~ Weight_mg, 
              data = data)
r_squared.all <- summary(model.all)$r.squared

# Calculate residuals
residuals.all <- residuals(model.all)

#good quality data
data.hq <- data[data$Failed_samples == "OK", ]


#### fit models ####
# model Spiked
model.hq <- lm(Final_pg.mg ~ Weight_mg, 
              data = data.hq)
r_squared.hq <- summary(model.hq)$r.squared

# Calculate residuals
residuals.hq <- residuals(model.hq)



# Mean Absolute Error
mae.all <- mean(abs(residuals.all))      
# Standard Deviation of Residuals
std_dev.all <- sd(residuals.all)

# Mean Absolute Error
mae.hq <- mean(abs(residuals.hq))      
# Standard Deviation of Residuals
std_dev.hq <- sd(residuals.hq)

```

**Error using good quality samples only**
```{r echo = FALSE}
cat("Number of good quality datapoints:", nrow(data.hq))
cat("Mean Absolute Error (MAE) ALL:", round(mae.hq, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.hq, 3), "\n")
```

**Error using all samples**
```{r echo = FALSE}
cat("Total number of datapoints:", nrow(data))
cat("Mean Absolute Error (MAE) ALL:", round(mae.all, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.all, 3), "\n")
```

