---
title: "QC_ELISA 3"
output: html_document
date: "2025-04-02"
---

Here I show how files are loaded, merged, and cleaned, including the exclusion of unnecessary columns and handling of missing values.

**Summary**

- Intra-Assay CV:  **11.35 %**

- Intra-Assay CV after removing low quality data: **5.799 %**

- Good quality data points: **30**

```{r define dataset}
# DATA SET 
current_test <- "Test3"
data_path <- file.path("./data", current_test)
```

# Set parameters

```{r}
# flag samples with high CV (15%) or binding above 80% and under 20%
CV_threshold <- 15.0
uppBinLim <- 80.0
lowBinLim <- 20.0
```


```{r libraries, echo = FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(dplyr)
```

# Data Cleaning and QC

Load, inspect and merge 3 files:

- layout: 7 columns (Wells, Sample, weight_mg, buffer_nl, spike, volume of the spike, dilution factor), 96 rows
- results: from myassays.com (not including standards)
- info: (Wells, Sample, Categ, TotalVol_well_uL)

```{r loading files, echo = FALSE}
# LAYOUT 

layout <- read.csv(file.path(data_path,"layout_wells_test3_091524.csv"), 
                   stringsAsFactors = TRUE, 
                   na.strings = c("", " "))

# RESULTS  

results <- read.csv(file.path(data_path,"myassays_table_test3_091524.csv"), 
                    stringsAsFactors = TRUE, 
                    na.strings = c("", " ", "-"))

results <- results[2:length(results)]


# INFO

info <- read.csv(file.path(data_path, "sample_info_test3_091524.csv"), 
                 stringsAsFactors = TRUE, 
                 na.strings = c("", " "))

```

Merged data set: 

```{r echo = FALSE}
# Merge files
m <- merge(layout, results, by = "Wells")
m <- merge(m, info, by = c("Sample", "Wells"))
m <- na.omit(m)
colnames(m) <- c("Sample", "Wells", "Raw.OD", 
                 "Binding.Perc", "Conc_pg.ml",
                 "Ave_Conc_pg.ml", "CV.Perc", 
                 "SD", "SEM", "Category", "Weight_mg", "Buffer_nl", 
                 "Spike", "TotalVol_well_ul", "SpikeVol_ul", "Dilution", "Sample_comparable") 
#m[, 3:11] <- lapply(m[, 3:11], 
 #                   function(x) as.numeric(as.character(x)))
m <- m[order(m$Sample),]
kable(head(m)) 

```

### Duplicates with high CV
Flag samples with high coefficient of variation (duplicate measurements that are too different from each other)
```{r echo = FALSE}
#create new file
m1 <- m %>%
  mutate(CV_categ = ifelse(CV.Perc > CV_threshold, 
                           "HIGH CV", NA))
cv_high <- m1 %>% 
  filter(CV.Perc > CV_threshold) %>% 
  arrange(desc(CV.Perc))

cat("High CV in a total of", nrow(cv_high), "duplicates. These are:")
print(cv_high[2])

```

### Samples outside the curve
Samples that have a binding percentage over 80 or 20 do not provide accurate results
```{r, echo = FALSE}
# flag samples with binding percentage over 80 or under 20
m2 <- m1 %>%
  mutate(Binding.Perc_categ = ifelse(Binding.Perc > uppBinLim, "ABOVE 80% binding", 
                                     ifelse(Binding.Perc < lowBinLim, "UNDER 20% binding", 
                                            NA)))

out_curve <- m2[!is.na(m2$Binding.Perc_categ), ]

cat(paste("Total samples outside the curve:", nrow(out_curve), sep=" "))
#kable(out_curve)

```

### Total samples failed

```{r echo = FALSE}
# dataset with failed samples flagged
data.flagged <- m2 %>%
  mutate(Failed_samples = ifelse(!is.na(CV_categ) & !is.na(Binding.Perc_categ), 
                             paste(CV_categ, Binding.Perc_categ, sep = ";"), 
                             coalesce(CV_categ, Binding.Perc_categ))) 

write.csv(data.flagged, file.path(data_path, "Data_QC_flagged.csv"), 
          row.names = FALSE) 

# table with samples that failed
failed_samples <- data.flagged[!is.na(data.flagged$Failed_samples),]
write.csv(failed_samples, file.path(data_path, "failed_samples.csv"), 
          row.names = FALSE) 

# data set with failed samples removed
data.no_failed <- m2 %>% 
 filter(CV.Perc < CV_threshold) %>%
 filter(Binding.Perc > lowBinLim & Binding.Perc < uppBinLim ) 
 
write.csv(data.no_failed, file.path(data_path, "Data_QC_filtered.csv"), row.names = FALSE) 

cat("Number of failed samples is", nrow(failed_samples))
cat("Number of good quality data points is", nrow(data.no_failed))

```

Location
```{r echo = FALSE}
cat("Data set with low quality samples flagged saved in: Data_QC_flagged.csv")
cat("Good quality data is stored in Data_QC_filtered.csv file")

```

Overall quality of the assay

```{r echo = FALSE}
CVraw <- mean(m2$CV.Perc)
CV <- mean(data.no_failed$CV.Perc)
cat("Intra-Assay CV including all samples is", CVraw, "%")
cat("Intra-Assay CV including only good quality data is", CV, "%")
```

