
---
title: "ELISA_computation"
author: "Paloma"
date: "2024-10-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

---

Here I calculate final cort concentration values using two methods. The first one (A) is the standard one, that does not correct for spike. The second one (B) does correct for spike on those samples that were spiked. 


```{r echo = FALSE, warning=FALSE, message=FALSE}
# Install libraries
library(dplyr)
library(knitr)
library(ggplot2)
```

```{r setup, include=TRUE, message=FALSE}
df <- read.csv("./data/Data_QC_flagged.csv")

# remove unnecessary information
data <- df %>% 
  select(!(c(Wells, Raw.OD, Concentration_pg.ml, SD, SEM))) %>%
  filter(CV.Perc < 15) %>% 
  filter(Binding.Perc < 80 & Binding.Perc > 20)

```

#A. Calculation

Formula:  ((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT concentration in pg/mg.


```{r}
kable(head(data))

# Transform to μg/dl from assay output
data$Average_Conc_ug.dL <- c(data$Average_Conc_pg.ml/10000)

# Creating variables in indicated units
data$Vol_methanol_ml <- 1.3/1

data$Buffer_ml <- c(data$Buffer_nl/1000)

# Calculate final values
data$final_conc_pg.mg <- c((data$Average_Conc_ug.dL / data$Weight_mg) * data$Vol_methanol_ml * data$Buffer_ml * 10000 )


```


## A. Results 
```{r}
# plot histogram

kable(head(data))
hist(data$final_conc_pg.mg, col = "cyan4", main = c("Histogram of cort concentration values, considering weight and dilutions"), xlab = "Cortisol concentrations, pg/mg", ylab = "Count", breaks = 60, xlim = range(0:60))

# scatterplot

ggplot(data, aes(y = final_conc_pg.mg, x = Weight_mg, color = as.factor(Spike))) +
  geom_point(size = 3) +  # Points with custom color and size
  geom_smooth(method = "lm", color = "gold3", se = FALSE) +  # Regression line without confidence interval
  theme_minimal() +  # Minimalistic theme
  labs(
    title = "Final Cort Concentration and Weight",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Center title and adjust size
    axis.title = element_text(size = 14),  # Adjust axis label size
    axis.text = element_text(size = 12)  # Adjust axis text size
  )

# summary for all samples

summary(data$final_conc_pg.mg)

```

#B. Recalculation (account for spike)

In **Nist et al. 2020**, they say "Thus, after pipetting 25μL of standards and samples into the appropriate wells of the 96-well assay plate, we added 25μL of the 0.333ug/dL standard to all samples,
resulting in a 1:2 dilution of samples. The remainder of the manufacturer’s protocol was
unchanged. We analyzed the assay plate in a Powerwave plate reader (BioTek, Winooski,
VT) at 450nm and subtracted background values from all assay wells. In the calculations, we
subtracted the 0.333ug/dL standard reading from the sample readings. Samples that resulted
in a **negative number were considered nondetectable**. We converted cortisol levels from
ug/dL, as measured by the assay, to pg/mg—based on the mass of hair collected and
analyzed using the following formula:

A/B * C/D * E * 10, 000 * 2 = F

where 
- A = μg/dl from assay output; 
- B = weight (in mg) of collected hair; 
- C = vol. (in ml) of methanol added to the powdered hair; 
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down; 
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract; 10,000 accounts for changes in metrics; 2 accounts for the dilution factor after addition of the spike; and 
- F = final value of hair cortisol concentration in pg/mg"


```{r}
dSpike <- data
#reading standard 1, in ug.dL
std.r <- ((3133 + 3146)/2)/10000

# Calculate final values
dSpike$final_conc_pg.mg <- ifelse(
  dSpike$Spike == 1,
    (dSpike$Average_Conc_ug.dL - std.r) / dSpike$Weight_mg * 1.3 / 1 * dSpike$Buffer_ml * 10000 * 2 ,
  dSpike$final_conc_pg.mg  # Keep existing value if Spike != 1
)

```

##B. Results 
```{r}

# histogram
hist(dSpike$final_conc_pg.mg, col = "cyan4", 
     main = c("Histogram of cort concentration values, considering weight, dilutions and spike"), 
     xlab = "Cortisol concentrations, pg/mg", 
     ylab = "Count", 
     breaks = 60)

# scatterplot
ggplot(dSpike, aes(y = final_conc_pg.mg, 
                   x = Weight_mg, 
                   color = as.factor(Spike))) +
  geom_point(size = 3) +  # Points with custom color and size
  geom_smooth(method = "lm", color = "gold3", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration and Weight, considering Spike",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"), 
    axis.title = element_text(size = 14),  # Adjust axis label size
    axis.text = element_text(size = 12)  # Adjust axis text size
  )

```

```{r}
# summary for all samples
summary(dSpike$final_conc_pg.mg)

# summary only for non-spiked samples
dSpikeSub <- dSpike[dSpike$Spike == 0,]
summary(dSpikeSub$final_conc_pg.mg)
```


<table>
<caption><span id="tab:table">Comparing results </span></caption>

Dilution:                 Nist et al.     My samples      Non-spiked only
---------------          -------------  -------------    -----------------
Mean cort conc (pg/mg)   23.74          -0.18              7.9 
Range cort conc (pg/mg)  2.1 to 124.9   -29.3 to 11.76     2.71 to 11.76
Weight range (mg)        0.4 to 10.9     11 to 37.1        12 to 37
Sample size              X              30                 18

</table>

* removed higher than 29 pg/mg as outliers


```{r echo = FALSE, include = FALSE}
# wflow_publish("./analysis/ELISA_Computation.Rmd")

```
