---
title: "Analysis of final values - test4"
author: "Paloma"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
# Introduction

Here I plan to include measurements and visualizations of the accuracy, precision, and linearity of ELISA #4. I will do this using two different ways of calculating final values: the standard method not accounting for spike, and an alternative method accounting for it. The way the spike is included in the formula is defined in file ELISA_Calc_FinalVals_test4.html.


```{r echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libraries, echo = FALSE, warning = FALSE, message=FALSE}
# Install libraries

library(knitr)
library(RColorBrewer)
library(stats)
library(coefplot)
library(arm)
library(bbmle)
library(plyr)
library(dplyr)
library(ggpmisc)
library(ggplot2)
```

```{r loading}
# loading info
# Loading data
std <- read.csv("./data/Test4/standard_data_test4.csv")
cort_valsA<- read.csv("./data/Test4/Data_cort_values_methodA.csv")
cort_valsB<- read.csv("./data/Test4/Data_cort_values_methodB.csv")
cort_valsD<- read.csv("./data/Test4/Data_cort_values_methodD.csv")
```

# Overall Results
## A) Standard method


```{r loading valsA}
# remove NSB
data <- cort_valsA[!(cort_valsA$Sample == "NSB"),]

#data <- data[,c("Sample", "Category", "Binding.Perc", "Final_conc_pg.mg", "Weight_mg", "Buffer_ml", "Spike", "Failed_samples", "Dilution")]
```

```{r expABCD.Final_conc_pg.mg_AD, fig.width=8, fig.height=6, results = 'hide'}
invisible({capture.output({ 
# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)


## PLOT COMPLETE DATASET ## 
ggplot(a, aes(x = Sample_num, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.05) +
  geom_smooth(formula = y ~ log(x), se = TRUE, linewidth = 0.5, linetype = 2, alpha = 0.1) +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 1) + 
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2, vjust = -1, hjust = 0.5, color = "gray30") +
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2.5) +
  labs(title = "Final vals not accounting for spike",
       subtitle = "Complete dataset",
       x = "Sample", y = "Final_conc_pg.mg Method A") +
  coord_cartesian(ylim = c(-5, 28)) +
  theme_minimal() 

exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "A2"))
})})
```

```{r expABCD.Final_conc_pg.mgA, fig.width=8, fig.height=6}

## PLOT WITHOUT POOR QUALITY SAMPLES ### 

ggplot(exp, aes(x = Sample_num, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.05) +
  geom_smooth(formula = y ~ log(x), se = TRUE, linewidth = 0.5, linetype = 2, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 1) + 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  labs(title = "Final vals not accounting for spike",
       subtitle = "removed high CV samples",
       x = "Sample", y = "Final_conc_pg.mg, Method A") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```

## B) Accounting for spike (Nist et al. 2020)

```{r load-B}
data <- cort_valsB[!(cort_valsB$Sample == "NSB"),]
```

```{r expABCD.Final_conc_pg.mgB, fig.width=8, fig.height=6, }

# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)
exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "A2"))

## PLOT WITHOUT POOR QUALITY SAMPLES ### 

ggplot(exp, aes(x = Sample_num, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.05) +
  geom_smooth(formula = y ~ log(x), se = TRUE, linewidth = 0.5, linetype = 2, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 1) + 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  labs(title = "Final vals accounting for spike",
       subtitle = "removed high CV samples",
       x = "Sample", y = "Final_conc_pg.mg, Method B") +
  scale_y_continuous(n.breaks = 10)  +
  theme_minimal() 
```

## D) New Method
```{r expABCD.Final_conc_pg.mgD, fig.width=8, fig.height=6}

data <- cort_valsD[!(cort_valsD$Sample == "NSB"),]

# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)

exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "A2"))

## PLOT WITHOUT POOR QUALITY SAMPLES ### 

ggplot(exp, aes(x = Sample_num, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.05) +
  geom_smooth(formula = y ~ log(x), se = TRUE, linewidth = 0.5, linetype = 2, alpha = 0.1) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 2.5, vjust = -1, hjust = 0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 1) + 
  stat_poly_eq(aes(x = Sample_num, label = paste(..rr.label.., sep = "~")),
               formula = y ~ x, parse = TRUE, size = 2) +
  labs(title = "Final vals accounting for spike",
       subtitle = "removed high CV samples",
       x = "Sample", y = "Final_conc_pg.mg, Method D") +
  scale_y_continuous(n.breaks = 10) + 
  theme_minimal() 

```


# Accuracy

Accuracy, or recovery %, is the ratio between observed and expected concentration, multiplied by 100. 
Expected values are defined using the dilution factor, starting with the value of the neat sample. 

- 80–120% is generally acceptable for biological assays.

- 70–130% may be acceptable at very low concentrations.

## Method A
```{r expABCD.conc.expect.vs.observed, results = 'hide', include = FALSE}

  ## Expected vals A
#a$Exp <- c(30.7800000, 8.655, 4.3275, 2.16375, 1.081875, 0.5409375, 0.2704687,
 #          34.2266667, 9.62625, 4.813125, 2.406563, 1.203281, 0.6016405, 0.3008203,
  #         16.4466667, 4.625625, 2.312813, 1.156406, 0.578203, 0.2891015, 0.1445508,
   #        17.193, 8.5965, 4.29825, 2.149125, 1.074563, 0.5372815, 0.2686407)
data <- cort_valsA[!(cort_valsA$Sample == "NSB"),]
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)

a$Expected <- NA
a$Expected[1:7] <- 30.7800000 / a$Dilution[1:7] 
a$Expected[8:14] <-  34.2266667/ a$Dilution[1:7] 
a$Expected[15:21] <-  16.4466667  / a$Dilution[1:7] 
a$Expected[22:28] <-  30.5653333 / a$Dilution[1:7]

ggplot(a, aes(x = Expected, 
                 y = (Final_conc_pg.mg), color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.1, linetype = 5) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray70") +
 geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gold") +
  geom_point(size = 2) +
  stat_poly_eq(aes(x = Expected, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 3) +

  labs(title = "Expected vs observed cortisol concentrations (not accounting for spike)",
       subtitle = "Complete dataset, final vals, method A",
       x = "Expected vals (pg/mL)", y = "Observed vals (pg/mL)") +
  theme_bw() + 
  facet_wrap(~ Category)

```

```{r expABCD.conc.expect.vs.observedA-filt}

exp <- a %>%
  dplyr::filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TD7", "TD6"))
tail(a,20)
ggplot(exp, aes(x = Expected, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.1, linetype = 5) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray50") +
  stat_poly_eq(aes(x = Expected, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Expected vs observed cortisol concentrations (not accounting for spike)",
       subtitle = "Removed low quality samples, final vals, methodA",
       x = "Expected vals (pg/mL)", y = "Observed vals (pg/mL)") +
  theme_bw() + 
  facet_wrap(~ Category)
``` 


## Method B

```{r loading valsB}
# remove NSB
data <- cort_valsB[!(cort_valsB$Sample == "NSB"),]
```

```{r expABCD.conc.expect.vs.observedB-filt, fig.width=8, fig.height=6}

# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)

a$Expected <- NA
a$Expected[1:7] <- 30.7800000 / a$Dilution[1:7] 
a$Expected[8:14] <-  25.6600000/ a$Dilution[1:7] 
a$Expected[15:21] <-   -9.9000000  / a$Dilution[1:7] 
a$Expected[22:28] <-  14.0580000 / a$Dilution[1:7]


## PLOT ## 
exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TD7", "TD6"))

ggplot(exp, aes(x = Expected, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.4, alpha = 0.1, linetype = 5) +
  geom_point(size = 2) +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = 0.5, color = "gray50") +
  stat_poly_eq(aes(x = Expected, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Expected vs observed cortisol concentrations (not accounting for spike)",
       subtitle = "Removed low quality samples, final vals, methodB",
       x = "Expected vals (pg/mL)", y = "Observed vals (pg/mL)") +
  theme_bw() + 
  facet_wrap(~ Category)


```

## Method D
```{r expABCD.conc.expect.vs.observed_D, include = FALSE, results = 'hide' }
# remove NSB
data <- cort_valsD[!(cort_valsD$Sample == "NSB"),]
## Expected vals B
#a$Exp <- c(17.31, 8.655, 4.3275, 2.16375, 1.081875, 0.5409375, 0.2704687,
#           25.0185000, 12.50925, 6.254625, 3.127312, 1.563656, 0.781828, 0.390914,
 #          -9.6525000, -4.82625, -2.413125, -1.206562,  -0.603281, -0.3016405,  -0.1508202,
  #         13.7065500, 8.5965, 4.29825, 2.149125, 1.074563, 0.5372815, 0.2686407)

# Scatter plot of Experiment A and B, observed binding % 
a <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Final_conc_pg.mg)) 
a$Sample_num <- 1:length(a$Sample)


a$Expected <- NA
a$Expected[1:7] <- 30.7800000 / a$Dilution[1:7] 
a$Expected[8:14] <-  32.2815152 / a$Dilution[1:7] 
a$Expected[15:21] <-  14.5015152  / a$Dilution[1:7] 
a$Expected[22:28] <-  18.7971667 / a$Dilution[1:7]

```

```{r expABCD.conc.expect.vs.observed_D-filt }

exp <- a %>%
  filter(!Sample %in% c("TA6","TA7","TB7","TC7", "TC6","TC5", "TD7", "TD6"))

ggplot(exp, aes(x = Expected, 
                 y = Final_conc_pg.mg, color = Category)) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.3, alpha = 0.1) +
 geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gold") +
  geom_point(size = 2) +
    geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 0.5) + 
  geom_text(aes(label = Sample), size = 2, vjust = -1, hjust = 0.5, color = "gray50") +
  stat_poly_eq(aes(x = Expected, label = paste(..rr.label.., sep = "~~~")),
               formula = y ~ x, parse = TRUE, size = 4) +
  labs(title = "All serial dilutions: expected vs observed cortisol concentrations",
       subtitle = "Removed low quality samples, final vals, methodD",
       x = "Expected vals (pg/mL)", y = "Observed vals (pg/mL)") +
  theme_bw() + 
  facet_wrap(~ Category)

``` 



# Precision

## Method A (controlling for dilution and weight)
```{r expE-concvals-final-vals-A}

# Scatter plot of Experiment D, observed binding % 

exp <- filter( cort_valsA, Category == "P", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length( exp$Sample )

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))

cv_values <- exp %>% 
  group_by(Group) %>%
  dplyr::summarise(
    cv_BP = round((sd( Final_conc_pg.mg ) / mean( Final_conc_pg.mg)) * 100, 1), 
    mean_BP = round( mean( Final_conc_pg.mg, na.rm = TRUE), 1),
    sd_BP = round(sd( Final_conc_pg.mg, na.rm = TRUE), 1)
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = round(mean( Final_conc_pg.mg, na.rm = TRUE),1),
    SD_BP = round(sd( Final_conc_pg.mg, na.rm = TRUE),1),
    CV_BP = round((SD_BP / Mean_BP) * 100, 1)
  )


ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_conc_pg.mg)) +
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
   labs(title = "Experiment E: Final conc vals (methodA) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_conc_pg.mg") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)

```

## Method B (controlling for dilution, weight and spike)
```{r expE-concvals-final-vals-B}

# Scatter plot of Experiment D, observed binding % 

exp <- filter(cort_valsB, Category == "P", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length(exp$Sample)

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))
cv_values <- exp %>%
  group_by(Group) %>%
  dplyr::summarise(
    CV_BP = round((sd( Final_conc_pg.mg) / mean(Final_conc_pg.mg)) * 100, 1), 
    mean_BP = mean( Final_conc_pg.mg, na.rm = TRUE),
    sd_BP = sd( Final_conc_pg.mg, na.rm = TRUE)
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = mean( Final_conc_pg.mg, na.rm = TRUE),
    SD_BP = sd( Final_conc_pg.mg, na.rm = TRUE),
    CV_BP = (SD_BP / Mean_BP) * 100
  )


ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_conc_pg.mg)) +
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", 
                  color = "red", linewidth = 1) +  
   labs(title = "Experiment E: Final conc vals (methodB) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_conc_pg.mg") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$CV_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)

```


## Method D (controlling for dilution and weight)
```{r expE-concvals-final-vals-D}

# Scatter plot of Experiment D, observed binding % 

exp <- filter( cort_valsD, Category == "P", !is.na(Binding.Perc)) 
exp$Sample_num <- 1:length( exp$Sample )

exp <- exp %>%
  mutate(Group = substr(Sample, 1, 3))

cv_values <- exp %>% 
  group_by(Group) %>%
  dplyr::summarise(
    mean_BP = round( mean( Final_conc_pg.mg, na.rm = TRUE), 1),
    sd_BP = round(sd( Final_conc_pg.mg, na.rm = TRUE), 1),
    cv_BP = round((sd( Final_conc_pg.mg ) / mean( Final_conc_pg.mg)) * 100, 1), 
  )

# overall stats
all <- exp %>%
  summarise(
    Mean_BP = round(mean( Final_conc_pg.mg, na.rm = TRUE),1),
    SD_BP = round(sd( Final_conc_pg.mg, na.rm = TRUE),1),
    CV_BP = round((SD_BP / Mean_BP) * 100, 1)
  )


ggplot(exp, aes(x = Weight_mg, 
                 y =  Final_conc_pg.mg)) +
geom_point(size = 3, color = "gray40", alpha = 1) +
  geom_pointrange(data = cv_values,
                  aes(x = c(6,9,12), y = mean_BP,
                      ymin = mean_BP - sd_BP, ymax = mean_BP + sd_BP),
                  color = "orange", size = 0.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.5, alpha = 0.05, color = "lightblue") +
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.5, color = "gray30") +
   labs(title = "Experiment E: Final conc vals (methodD) by sample weight",
       subtitle = "Precision, 25uL spike, 25uL sample",
       x = "Sample weight (mg)", y = " Final_conc_pg.mg") +
  scale_y_continuous(n.breaks = 10) +
  coord_cartesian(xlim = c(4,14)) +
  theme_minimal() +
  annotate("text", x = 6, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP1"], "%"), size = 4) +
  annotate("text", x = 9, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP2"], "%"), size = 4) +
  annotate("text", x = 12, y = 35, label = paste0("CV: ", cv_values$cv_BP[cv_values$Group == "TP3"], "%"), size = 4) +
  annotate("text", x = 6, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP1"],2)), size = 4) +
  annotate("text", x = 9, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP2"],2)), size = 4) +
  annotate("text", x = 12, y = 40, label = paste0("mean: ", round(cv_values$mean_BP[cv_values$Group == "TP3"],2)), size = 4) 

print(cv_values)

```
