---
title: "Cortisol Concentration Values, Test4"
author: "Paloma Contreras"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Summary

**Cortisol value calculations** 

|                        | Min.   | 1st Qu. |  Median |  Mean   | 3rd Qu. | Max.   |  NA's |
|------------------------|--------|---------|---------|---------|---------|--------|-------|
|**A) Standard Method**| 0.4142 | 1.9653  | 7.5533  | 14.1906 | 28.0667 | 50.933 |    3  |
|**B) Spike-Corrected Method**|-45.870 | -31.922 | -1.929 | -9.444 |  7.553 | 30.780  |  3 |
|**C) Standard, but simplified equation**| 0.4142 |1.9653 | 7.5533 | 14.1906 | 28.0667| 50.9333 |  3 |
|**D) Spike-Corrected (divided by two)**|  0.2335 | 1.5067  |6.7184|  9.7699| 17.3683 |32.2815    |   3 |



**Results**:  


- Intra-assay CV: 14.5% 

- Intra-assay CV after removing low quality samples: 10%

- Inter-assay CV: 21% 

(Bindings for 20mg sample diluted in 250 uL, no spike: 64.8% and 48% in test3 and test4, respectively)


**Conclusions**: 




**Concerns:**
Overall quality of the plate is not great, but serial dilusions show clear parallelism and standards have values within the expected


# Cortisol concentration calculations

```{r parameters}
# define volume of methanol used for cortisol extraction
# vol added / vol recovered (mL)
extraction <- 1/0.750

# set reading value of spike (std1, 0.333 ug/dL), 
# and transforming to ug.dL

std <- (3191+3228)/2
std.r <- (std/10000)
std
std.r

# according to chatgpt, the spike's contribution is
# 1600 pg/mL, which is very similar to half of the reading for std 1 :]
```

```{r echo = FALSE, message=FALSE}
#path:
data_path = "./data/Test4"
library(knitr)
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
```

Loading files and transforming units, *including low quality data*

```{r setup, include=TRUE, message=FALSE}
df <- read.csv(file.path(data_path,"Data_QC_flagged.csv"))
kable(tail(df))
# remove outlier
#df<- df[(df$Sample != "TP3A"),]

# Creating variables in indicated units
# dilution (buffer)
df$Buffer_ml <- c(df$Buffer_nl/1000)

# remove unnecessary information 
data <- df %>%
    dplyr::select(Wells, Sample, Category, Binding.Perc, Ave_Conc_pg.ml, Weight_mg, Buffer_ml, Spike, SpikeVol_uL, Dilution, Vol_in_well.tube_uL, Failed_samples) 

kable(tail(data, 10))
dim(data)

# remove duplicates
data.og <- data[!is.na(data$Binding.Perc), ]

```

## (A) Standard Calculation

Formula: 

((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r A}
##################################
##### Calculate final values #####
##################################

# Transform to μg/dl from assay output
data$Ave_Conc_ug.dL <- c(data$Ave_Conc_pg.ml/10000)

data$Final_conc_pg.mg <- c(
    ((data$Ave_Conc_ug.dL) / data$Weight_mg) * # A/B *
      extraction *                                  # C/D  *     
      data$Buffer_ml * 10000 )                 # E * 10000
data <- data[order(data$Sample),]
write.csv(data, file.path(data_path, "Data_cort_values_methodA.csv"), row.names = F)

# summary for all samples
summary(data$Final_conc_pg.mg)

kable(tail(data, 7))
dim(data)

```

## (B) Accounting for Spike

We followed the procedure described in **Nist et al. 2020**:

"Thus, after pipetting 25μL of standards and samples into the appropriate wells of the 96-well assay plate, we added 25μL of the 0.333ug/dL standard to all samples,
resulting in a 1:2 dilution of samples. The remainder of the manufacturer’s protocol was
unchanged. We analyzed the assay plate in a Powerwave plate reader (BioTek, Winooski,
VT) at 450nm and subtracted background values from all assay wells. In the calculations, we
subtracted the 0.333ug/dL standard reading from the sample readings. Samples that resulted
in a **negative number were considered nondetectable**. We converted cortisol levels from
ug/dL, as measured by the assay, to pg/mg—based on the mass of hair collected and
analyzed using the following formula:

A/B * C/D * E * 10,000 * 2 = F

where 
- A = μg/dl from assay output; 
- B = weight (in mg) of collected hair; 
- C = vol. (in ml) of methanol added to the powdered hair; 
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down; 
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract; 10,000 accounts for changes in metrics; 2 accounts for the dilution factor after addition of the spike; and 
- F = final value of hair cortisol concentration in pg/mg"


```{r B-calc}
dSpike <- data

##################################
##### Calculate final values #####
##################################

dSpike$Final_conc_pg.mg <- 
  ifelse(
    dSpike$Spike == 1,    ## Only spiked samples
      ((dSpike$Ave_Conc_ug.dL - (std.r)) / # (A-spike) / B
        dSpike$Weight_mg) 
        * extraction *                  # C / D
        dSpike$Buffer_ml * 10000 * 2,    # E * 10000 * 2
        dSpike$Final_conc_pg.mg  
    )


write.csv(dSpike, file.path(data_path, "Data_cort_values_methodB.csv"), row.names = F)

# summary for all samples
summary(dSpike$Final_conc_pg.mg)


dSpikeSub <- data[c(data$Spike == 0), ]
summary(dSpikeSub$Final_conc_pg.mg)

kable(tail(dSpike, 10))
```

## (C) Skip unit transformation


```{r C-calc}
##################################
##### Calculate final values #####
################################## 
datac <- data
datac$Final_conc_pg.mg <- c(
    (datac$Ave_Conc_pg.ml / datac$Weight_mg) * # A/B *
      extraction *                                  # C/D  *     
      datac$Buffer_ml)                 # E 
datac <- datac[order(datac$Sample),]
write.csv(datac, file.path(data_path, "Data_cort_values_methodC.csv"), row.names = F)

# summary for all samples
summary(datac$Final_conc_pg.mg)

```

## (D) Account for Spike contribution (uses vol. of spike, conc. of spike, and total reconstit. vol.)

Spike contribution (pg/mL) = (Vol. spike (mL) x Conc. spike (pg/mL) ) / Vol. reconstitution (mL) or total vol. in well (50uL) (depending on where the spike was added)
```{r D-calc}
# Calculate contribution of spike according to the different volumes in which it was added
# Consider that contribution of spike in serial dilutions gets smaller 

# Vol. of spike transformed to mL
data$SpikeVol_ml <- data$SpikeVol_uL/1000
# Concentration of the spike:
std
# Vol. reconstitution (mL) is the total volume in tube or well (sample + spike), after adding spike.
# transform to mL
data$Vol_in_well.tube_ml <- data$Vol_in_well.tube_uL/1000


  ##( Spike vol. x Spike Conc.)
  ## ------------------------  / dilution = Spike contribution
  ##      Total vol. 
  
# Cortisol added by spike in wells: 0.0025 mL x 3200 pg/mL = 80 pg
# Calculate cort contribution of spike to each sample
data$Spike.cont_pg.mL <- (((data$SpikeVol_ml * std ) / # Volume of spike * Spike concentration
                            data$Vol_in_well.tube_ml) / # divided by the total volume (spike + sample)
                              data$Dilution) # resulting number changes depending on the dilution

dSpiked <- data
                          
##################################
##### Calculate final values #####
##################################


dSpiked$Final_conc_pg.mg <- 
  ifelse(
    dSpiked$Spike == 1,           ## Only spiked samples
      ((dSpiked$Ave_Conc_pg.ml - dSpiked$Spike.cont_pg.mL) / # (A - spike) / B
      dSpiked$Weight_mg) 
    * extraction *      # C / D
      dSpiked$Buffer_ml,    # E * 
    dSpiked$Final_conc_pg.mg  
)

write.csv(dSpiked, file.path(data_path, "Data_cort_values_methodD.csv"), row.names = F)

# summary for all samples
summary(dSpiked$Final_conc_pg.mg)
kable(tail(dSpiked[!is.na(dSpiked$Final_conc_pg.mg) , c("Sample", "Final_conc_pg.mg", "Ave_Conc_pg.ml", "Spike.cont_pg.mL", "Binding.Perc", "Weight_mg", "Buffer_ml", "SpikeVol_uL", "Dilution", "Vol_in_well.tube_uL")],20))
```

# Plots 

## (A) Standard Calculation

```{r echo = FALSE, warning=FALSE }
# scatterplot method A
data$Spike <- replace(data$Spike, data$Spike == 1, 'Yes')
data$Spike <- replace(data$Spike, data$Spike == 0, 'No')
data$Buffer <- data$Buffer_ml
data$Buffer <- replace(data$Buffer, data$Buffer == 0.06, '60 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.11, '110 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.22, '220 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.25, '250 uL')


ggplot(data, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 2.5, alpha = 0.85) +  
   geom_text(aes(label = Sample), size = 2.5, vjust = -0.65, hjust = -0.18) +
    theme_minimal() +  
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  xlim(0,52) +
  labs(
    title = "(A) Standard Calculation Cortisol Values",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)") + 
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  ) + 
  scale_color_paletteer_d("vangogh::CafeTerrace")
```

## (B) Accounting for Spike
```{r echo = FALSE, warning=FALSE }

dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 1, 'Yes')
dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 0, 'No')
dSpike$Buffer <- dSpike$Buffer_ml
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.06, '60 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.11, '110 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.22, '220 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.25, '250 uL')

# scatterplot
ggplot(dSpike, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 3.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
  xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(B) Calculation Accounting for Spike",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )+ 
  scale_color_paletteer_d("vangogh::CafeTerrace")

```

## (C) Simplified Standard Calculation
```{r echo = FALSE, warning=FALSE }
# scatterplot method c
datac$Spike <- replace(datac$Spike, datac$Spike == 1, 'Yes')
datac$Spike <- replace(datac$Spike, datac$Spike == 0, 'No')
datac$Buffer <- datac$Buffer_ml
datac$Buffer <- replace(datac$Buffer, datac$Buffer == 0.06, '60 uL')
datac$Buffer <- replace(datac$Buffer, datac$Buffer == 0.11, '110 uL')
datac$Buffer <- replace(datac$Buffer, datac$Buffer == 0.22, '220 uL')
datac$Buffer <- replace(datac$Buffer, datac$Buffer == 0.25, '250 uL')


ggplot(datac, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 2.5, alpha = 0.85) +  
   geom_text(aes(label = Sample), size = 2.5, vjust = -0.65, hjust = -0.18) +
    theme_minimal() +  
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
    #ylim(-26,24) +
#  xlim(0,52) +
  labs(
    title = "(C) ",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)") + 
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  ) + 
  scale_color_paletteer_d("vangogh::CafeTerrace")
```

## (D) New calculation (substract half of the spike from A)
```{r echo = FALSE, warning=FALSE }

dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 1, 'Yes')
dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 0, 'No')
dSpiked$Buffer <- dSpiked$Buffer_ml
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.06, '60 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.11, '110 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.110, '110 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.25, '250 uL')

# scatterplot
ggplot(dSpiked, aes(y = Final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = Spike,
                 shape = Buffer)) +
  geom_point(size = 3.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
    ylim(-26,30) +
  xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(D) ",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )+ 
  scale_color_paletteer_d("vangogh::CafeTerrace")

```
