---
title: "Calculation Final Values, Test7"
author: "Paloma Contreras"
date: "`r Sys.Date()`"
output: workflowr::wflow_html

---

# Summary

**Cortisol value calculations** 


|                        | Min. | 1st Qu. | Median | Mean | 3rd Qu. | Max. | NA's |
|------------------------|------|---------|---------|---------|---------|-------|------|
|**A) Standard Method (mult. by sample dilution)**| |
|**B) Spike-Corrected Method (Nist 2020)**|  |
|**C) Spike-Corrected (Sam's Method)**|  |


*Plate description*

Category	  Description
---------  ---------------------------------------------------------------




**Results:**  

- Intra-assay CV: -- % 

- Intra-assay CV after removing low quality samples: --%

- Inter-assay CV: **-- ** (Bindings for 20mg sample diluted in 250 uL, no spike: 64.8% and 48% in test3 and test4, respectively)


**Conclusions**: 



**Concerns:**


## Explanation of each variable used in calculations

- Ave_Conc_pg/ml: average ELISA reading per sample in pg/mL

- Weight_mg: hair weight in mg

- Buffer_nl: assay buffer volume in nL → we convert to mL

- Dilution: dilution factor (already included in arborassays.com)

- std: standard reading value

- extraction: methanol volume ratio = vol added / vol recovered (e.g. 1.4/1 ml)


# Cortisol concentration calculations

```{r Loading files, echo = FALSE, message=FALSE}
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
library(knitr)
library(here)

#path:
data_path <- "./data/Test7"
print(file.path(data_path, "Data_QC_flagged.csv"))  # Check the full path
df <- read.csv(file.path(data_path, "Data_QC_flagged.csv"))
#df <- read.csv(here("data", "Test6", "Data_QC_flagged.csv"))


```

Parameters and unit transformations: 
```{r parameters}
# Volume of methanol used for cortisol extraction varies, so it is included in file
# as Extraction_ratio (vol added / vol recovered) in mL

# Reading of spike standard and conversion to ug/dl
std <- 3200                                  # test 6 backfit [FAKE]
std_ul.dl <- std / 10000                        # std in ul/dl

# Creating variables in indicated units
df$Buffer_ml <- c(df$Buffer_nl/1000)            # dilution (buffer)
df$Ave_Conc_ug.dl <- c(df$Ave_Conc_pg.ml/10000) # Transform to μg/dl from assay output
```

Identify and flag bad quality samples
```{r filter, include=TRUE, message=FALSE, echo = FALSE}

# remove outliers

# label samples by quality

df$Failed_samples[is.na(df$Failed_samples)] <- "OK"
table(df$Failed_samples)
df$Failed_samples[df$Failed_samples == "ABOVE 80% binding"] <- "Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;ABOVE 80% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;UNDER 20% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV"] <- "High CV"
df$Failed_samples[df$Failed_samples == "UNDER 20% binding"] <- "Out of curve"

df$Failed_samples <- factor(
  df$Failed_samples,
  levels = c(
    "OK",
    "Out of curve",
    "High CV",
    "High CV & Out of curve"
  )
)

cat("Summary CV for all samples:")
summary(df$CV.Perc)

cat("Summary CV for good quality samples only:")
temp <- df %>% 
  filter(Failed_samples == 
           "OK")
summary(temp$CV.Perc)


# remove unnecessary information 
data <- df %>%
    dplyr::select(Wells, Sample, Category, Binding.Perc, 
                  Ave_Conc_ug.dl, Ave_Conc_pg.ml, Weight_mg, 
                  Buffer_ml, Dilution_sample, 
                 Extraction_ratio,  
                  Failed_samples) 



data <- data[!is.na(data$Binding.Perc), ] # remove duplicates

```

## (A) Standard Calculation

Formula: 

((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg

```{r A}
##################################
##### Calculate final values #####
##################################

data$Final_pg.mg_A <- c(
    ((data$Ave_Conc_ug.dl / data$Weight_mg) *              # A/B *
      data$Extraction_ratio *                              # C/D  *     
      data$Buffer_ml * 10000))                             # E * 10000 

```

```{r echo=FALSE}
data <- data[order(data$Sample),]
write.csv(data, file.path(data_path, "Data_cort_values_methodA.csv"), row.names = F)

# summary for all samples
data <- data %>% 
  filter(!Sample %in% c("B0", "BE", "POOL")) 

cat("Summary of all samples (n =", nrow(data), "):")
summary(data$Final_pg.mg_A)

temp <- data %>% 
  filter(Failed_samples == 
           "OK")
cat("Summary for good quality samples only (n =", nrow(temp), "):")

summary(temp$Final_pg.mg_A)
```

## (C) Sam's calculation

Simplifies unnecessary unit transformations and accounts for spike considering dilution of both sample and the spike


Calculate final values accounting for dilution of the sample, weight, and reconstitution


((A)/B) * (C/D) * E  = F

- A = pg/ml from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r}
##################################
##### Calculate final values #####
##################################

data$Final_pg.mg_C <- 
     (data$Ave_Conc_pg.ml / data$Weight_mg) *         # A / B *
     data$Extraction_ratio *                          # C / D *
     data$Buffer_ml                                   # E 
```

   


```{r echo = FALSE}
write.csv(data, file.path(data_path, "Data_cort_values_methodC.csv"), row.names = F)

# summary for all samples
data <- data %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL")) 
cat("Summary for all samples:")
summary(data$Final_pg.mg_C)

cat("Summary for good quality samples only:")
temp <- data %>% 
  filter(Failed_samples == 
           "OK")
summary(temp$Final_pg.mg_C)

kable(tail(data[ , c("Sample", "Final_pg.mg_A", "Final_pg.mg_B", "Final_pg.mg_C", "Binding.Perc", "Dilution_sample", "Extraction_ratio")]), n= 3)

```

# Plots: all samples vs final concentration

## (A) Standard Calculation

Tags are sample numbers.

Expected results:

```{r echo = FALSE, warning=FALSE }
# scatterplot method A

data$Buffer <- data$Buffer_ml

data$Buffer <- replace(data$Buffer, data$Buffer == 0.22, '220 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.11, '110 uL')

ggplot(data, aes(y = Final_pg.mg_C, 
                 x = 1:length(Sample), 
                 color = Failed_samples,
                 shape = Buffer)) +
  geom_point(size = 2.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
 #    ylim(-26,30) +
  #   xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(C) Sam's calculation",
    y = "Final Concentration (pg/mg)",
    x = "Sample" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  )+ 
  scale_color_paletteer_d("vangogh::CafeTerrace")
```


```{r echo = FALSE, warning=FALSE, message = FALSE, include = FALSE}
# scatterplot
#dataplot <- filter(data, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Binding.Perc)) 

#ggplot(dataplot, aes(y = Final_pg.mg_C, 
 #                x = 1:length(dataplot$Sample), 
    #             color = Failed_samples,
     #            shape = Buffer)) +
#  geom_point(size = 2.5,  alpha = 0.85) +  
#  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
#  theme_minimal() +  
 #   ylim(-26,30) +
  #xlim(0,52) +
 # geom_hline(yintercept = 0, 
  #           linetype = "dashed", color = "red") +
#  labs(
 #   title = "(C) Sam's calculation",
#    y = "Final Concentration (pg/mg)",
#    x = "Sample" ) +
#  theme(
#    plot.title = element_text(hjust = 0.5, 
 #                             size = 17, face = "bold"),
  #  axis.title = element_text(size = 14),  
   # axis.text = element_text(size = 10)  
#  )+ 
 # scale_color_paletteer_d("vangogh::CafeTerrace")
```

# Plot grouped by category (C)
```{r plot-by-categ, echo = FALSE}
ggplot(data2, aes(y = Final_pg.mg_C, 
                  x = 1:length(Sample),
                  fill= Category)) +
  geom_point(size = 2.5) +  
  geom_text(label = c(data2$Sample), nudge_y = 0.75, nudge_x = -1, size = 2) +
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.3) + 
  geom_hline(yintercept = mean(data2$Final_pg.mg_C), 
             color = "lightblue3",
             linetype = "dashed") +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration by Category
    (All samples) method C",
    y = "Final Concentration (pg/mg)",
    x = "Sample number"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) 
```


# Error evaluations, grouped by data quality (using C)
```{r models-good-quality, message=FALSE, echo = FALSE}

#good quality data
data2.hq <- data2[data2$Failed_samples == "OK", ]
cat("Number of good quality datapoints:", nrow(data2.hq))

cat("Models used with respect to weight: 

      lm(Final.pg.mg_C ~ Weight_mg, 
                  data = data2.hq)")
#### fit models ####
# model Spiked
model.hq <- lm(Final_pg.mg_C ~ Weight_mg, 
              data = data2.hq)
r_squared.hq <- summary(model.hq)$r.squared

# Calculate residuals
residuals.hq <- residuals(model.hq)

# Quantify residuals

# Mean Absolute Error
mae.all <- mean(abs(residuals.all))      
# Standard Deviation of Residuals
std_dev.all <- sd(residuals.all)

# Mean Absolute Error
mae.hq <- mean(abs(residuals.hq))      
# Standard Deviation of Residuals
std_dev.hq <- sd(residuals.hq)

```

**Error using good quality samples only**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) ALL:", round(mae.hq, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.hq, 3), "\n")
```

**Error using all samples**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) ALL:", round(mae.all, 3), "\n") 
cat("Standard Deviation of Residuals ALL:", round(std_dev.all, 3), "\n")
```

