---
title: "ELISA_computation"
author: "Paloma"
date: "2024-10-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Summary

**Data Cleaning and Quality Control (QC)** details how data files are loaded, merged, and cleaned, including the exclusion of unnecessary columns and handling of missing values.

**Cortisol value calculations** were conducted using two methods:

- **Standard Method (Method A)**: Calculates cortisol concentration without correction for spiked samples.
- **Spike-Corrected Method (Method B)**: Adjusts for spiked samples to account for additional cortisol added for testing accuracy.

**Results**: The formula to account for the spike added to some samples results in final cortisol concentration values become negative, which would mean that there is no cortisol in original samples. We believe that this is an artifact of an extremely high absorbance level. Non-spiked samples, however, result in values that are within the range found in other studies of cortisol in human hair. After accounting for differences in dilution and weight, our results suggest some optimal parameters. 

**Conclusions**: 

- Spiking samples does not provide consistent results
- Dilution of 250uL is preferable over 60uL 
- Use of samples with a mass of 20mg or higher is recommended


```{r echo = FALSE, warning=FALSE, message=FALSE}
# Install libraries
library(dplyr)
library(knitr)
library(ggplot2)
library(broom)
```

# Data Cleaning and QC


```{r}
# Load files
layout <- read.csv("./data/layout_wells_test3_091524.csv", stringsAsFactors = TRUE, na.strings = c("", " "))
results <- read.csv("./data/myassays_table_test3_091524.csv", stringsAsFactors = TRUE, na.strings = c("", " ", "-"))
info <- read.csv("./data/sample_info_test3_091524.csv", stringsAsFactors = TRUE, na.strings = c("", " "))

#kable(summary(c(layout, results, info)))

```


```{r}

# Merge and clean files
kable(head(layout)) 
kable(head(results)) 
dim(results)
results <- results %>% 
  select(!Sample) 

m <- merge(layout, results, by = "Wells")
kable(head(m)) 
dim(m)

m <- merge(m, info, by = c("Sample", "Wells"))
m <- na.omit(m)
kable(head(m)) 
dim(m)


colnames(m) <- c("Sample", "Wells", "Raw.OD", "Binding.Perc", "Concentration_pg/ml","Average_Conc_pg/ml", "CV.Perc", "SD", "SEM", "Weight_mg", "Buffer_nl", "Spike") 

m[, 3:11] <- lapply(m[, 3:11], function(x) as.numeric(as.character(x)))
m$Spike <- as.factor(m$Spike)


```



### Duplicates with high CV

```{r}
# define high Coef. of variation as higher than 15%
CV_threshold <- 15.0
cv_high <- m %>% 
  filter(CV.Perc > CV_threshold) %>% 
  arrange(desc(CV.Perc))

kable(cv_high)

# flag high CV

m1 <- m %>%
  mutate(CV_categ = ifelse(CV.Perc > CV_threshold, "HIGH CV", NA))

```

### Samples outside the curve

```{r}

out_curve <- m %>% 
  filter(Binding.Perc < 20.0 | Binding.Perc > 80.0) %>% 
  arrange(desc(Binding.Perc))

kable(out_curve)


# flag samples outside the curve

m2 <- m1 %>%
  mutate(Binding.Perc_categ = ifelse(Binding.Perc > 80, "ABOVE 80% binding", ifelse(Binding.Perc < 20, "UNDER 20% binding", NA)))

m2 <- m2 %>%
  mutate(Failed_samples = ifelse(!is.na(CV_categ) & !is.na(Binding.Perc_categ), 
                             paste(CV_categ, Binding.Perc_categ, sep = ";"), 
                             coalesce(CV_categ, Binding.Perc_categ))) %>%
  select(!c(CV_categ, Binding.Perc_categ))

# dataset with problematic samples flagged

kable(head(m2))
write.csv(m2, "./data/Data_QC_flagged.csv", row.names = FALSE) 


# dataset with high cv samples removed
data.no_cv <- m2 %>% 
 filter(CV.Perc < CV_threshold) 


write.csv(data.no_cv, "./data/Data_QC_filtered.csv", row.names = FALSE) 

# table with samples that failed
failed_samples <- rbind(cv_high, out_curve)
write.csv(failed_samples, "./data/failed_samples.csv", row.names = FALSE) 

```


```{r setup, include=TRUE, message=FALSE}
df <- read.csv(
  "./data/Data_QC_flagged.csv"
  )

# remove unnecessary information
data <- df %>% 
  select(-Wells, -Raw.OD, -Concentration_pg.ml, -SD, -SEM) %>%
  filter(CV.Perc < 15) %>% 
  filter(Binding.Perc < 80 & Binding.Perc > 20)

```

# (A) Standard Calculation

Formula: 

((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT concentration in pg/mg.


```{r}
kable(head(data))

# Transform to μg/dl from assay output
data$Average_Conc_ug.dL <- c(data$Average_Conc_pg.ml/10000)

# Creating variables in indicated units
data$Vol_methanol_ml <- 1.3/1

data$Buffer_ml <- c(data$Buffer_nl/1000)

# Calculate final values
data$final_conc_pg.mg <- c(
    (data$Average_Conc_ug.dL / data$Weight_mg) *
      data$Vol_methanol_ml * 
      data$Buffer_ml * 10000)


```


## Results 
```{r}
kable(head(data[ , c("Sample",
                     "Binding.Perc", 
                     "Vol_methanol_ml", 
                     "Buffer_ml",
                     "final_conc_pg.mg")]))

# plot histogram

hist(data$final_conc_pg.mg, 
     col = "cyan4", 
     main = c("Histogram of final cort concentration values"), 
     xlab = "Cortisol concentrations, pg/mg", 
     ylab = "Count", 
     breaks = 60, 
     xlim = range(0:60))

# scatterplot

ggplot(data, aes(y = final_conc_pg.mg, 
                 x = Weight_mg, 
                 color = as.factor(Spike))) +
  geom_point(size = 3) +  
  geom_smooth(method = "lm", 
              color = "gold3", se = FALSE) +  
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration and Weight",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )

# summary for all samples

summary(data$final_conc_pg.mg)

```

# (B) Calculation accounting for spike

We followed the procedure described in **Nist et al. 2020**:

"Thus, after pipetting 25μL of standards and samples into the appropriate wells of the 96-well assay plate, we added 25μL of the 0.333ug/dL standard to all samples,
resulting in a 1:2 dilution of samples. The remainder of the manufacturer’s protocol was
unchanged. We analyzed the assay plate in a Powerwave plate reader (BioTek, Winooski,
VT) at 450nm and subtracted background values from all assay wells. In the calculations, we
subtracted the 0.333ug/dL standard reading from the sample readings. Samples that resulted
in a **negative number were considered nondetectable**. We converted cortisol levels from
ug/dL, as measured by the assay, to pg/mg—based on the mass of hair collected and
analyzed using the following formula:

A/B * C/D * E * 10, 000 * 2 = F

where 
- A = μg/dl from assay output; 
- B = weight (in mg) of collected hair; 
- C = vol. (in ml) of methanol added to the powdered hair; 
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down; 
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract; 10,000 accounts for changes in metrics; 2 accounts for the dilution factor after addition of the spike; and 
- F = final value of hair cortisol concentration in pg/mg"


```{r}
dSpike <- data
#reading standard 1, in ug.dL
std.r <- ((3133 + 3146)/2)/10000

# Calculate final values
dSpike$final_conc_pg.mg <- 
  ifelse(
    dSpike$Spike == 1,
      (dSpike$Average_Conc_ug.dL - std.r) / 
      dSpike$Weight_mg * 1.3 / 1 * 
      dSpike$Buffer_ml * 10000 * 2 ,
    dSpike$final_conc_pg.mg  # Keep value if Spike != 1
)

```

## Results 
```{r}

# histogram
hist(dSpike$final_conc_pg.mg, 
     col = "cyan4", 
     main = c("Final cort values, spike adjusted
              as in Nist et al. 2020"), 
     xlab = "Cortisol concentrations, pg/mg", 
     ylab = "Count", 
     breaks = 60)

# scatterplot
ggplot(dSpike, aes(y = final_conc_pg.mg, 
                   x = Weight_mg, 
                   color = as.factor(Spike))) +
  geom_point(size = 3) +  
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", 
             color = "red") +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration and Weight,
    adjusted for Spike",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 16, 
                              face = "bold"), 
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12)  
  )

```

```{r}
# summary for all samples
summary(dSpike$final_conc_pg.mg)

# summary only for non-spiked samples
dSpikeSub <- dSpike [ dSpike$Spike == 0, ]
summary(dSpikeSub$final_conc_pg.mg)
```

# Comparing A and B

<table>
<caption><span id="tab:table">Comparing results </span></caption>

Dilution:                 Nist et al.     My samples      Non-spiked only
---------------          -------------  -------------    -----------------
Mean cort conc (pg/mg)   23.74          -0.18              7.9 
Range cort conc (pg/mg)  2.1 to 124.9   -29.3 to 11.76     2.71 to 11.76
Weight range (mg)        0.4 to 10.9     11 to 37.1        12 to 37
Sample size              X              30                 18

</table>

Consider that Nist et al. 2020 removed higher than 29 pg/mg as outliers


# Cortisol concentrations, non-spiked samples only

Since all spiked samples produce negative values, we will continue our analyses using only non-spiked samples, which produced values more similar to what is expected from hair cortisol in humans.

```{r}
# non-spiked samples only
data2 <-data[data$Spike == 0, ]

#two datasets, separated by dilution
data2.06 <- data2[data2$Buffer_ml == 0.06, ]
data2.25 <- data2[data2$Buffer_ml == 0.25, ]

kable(head(data2[ , c("Sample",
                     "Binding.Perc", 
                     "Vol_methanol_ml", 
                     "Buffer_ml",
                     "final_conc_pg.mg")]))
# plot histogram
hist(data2$final_conc_pg.mg, 
     col = "cyan4", 
     main = c("Final cort concentration 
              values, non-Spiked only"), 
     xlab = "Cortisol concentrations, pg/mg", 
     ylab = "Count", 
     breaks = 60)


# fit models 
model <- lm(final_conc_pg.mg ~ Weight_mg, 
            data = data2)

# model Buffer = 0.06
model06 <- lm(final_conc_pg.mg ~ Weight_mg, 
              data = data2[data2$Buffer_ml == 0.06,])
r_squared06 <- summary(model06)$r.squared

# model Buffer = 0.25
model25 <- lm(final_conc_pg.mg ~ Weight_mg, 
              data = data2[data2$Buffer_ml == 0.25,])
r_squared25 <- summary(model25)$r.squared

# Calculate residuals
residuals06 <- residuals(model06)
residuals25 <- residuals(model25)

# Quantify residuals
# Mean Absolute Error
mae06 <- mean(abs(residuals06))          
# Standard Deviation of Residuals
std_dev06 <- sd(residuals06)   

# Mean Absolute Error
mae25 <- mean(abs(residuals25))      
# Standard Deviation of Residuals
std_dev25 <- sd(residuals25)                      

# scatterplot

ggplot(data2, aes(y = final_conc_pg.mg, 
                  x = Weight_mg, 
                  color = as.factor(Buffer_ml), 
                  fill = as.factor(Buffer_ml))) +
  geom_point(size = 3) +  
  geom_smooth(method = "lm", 
              color = "gold3", 
              se = TRUE,
              alpha = 0.1) + 
  geom_hline(yintercept = mean(data2$final_conc_pg.mg), 
             color = "gray90",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.06$final_conc_pg.mg), 
             color = "lightpink2",
             linetype = "dashed") +
  geom_hline(yintercept = mean(data2.25$final_conc_pg.mg), 
             color = "lightblue3",
             linetype = "dashed") +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration and Weight
    (Non-spiked only)",
    y = "Final Concentration (pg/mg)",
    x = "Weight (mg)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) +
  # Add R^2 annotation
  annotate("text", x = max(data2$Weight_mg) * 0.7, y = min(data2$final_conc_pg.mg) * 1.5,
           label = paste("R² =", round(r_squared06, 3)), size = 5, color = "black") +
  annotate("text", x = max(data2$Weight_mg) * 0.7, y = max(data2$final_conc_pg.mg) * 0.84,
           label = paste("R² =", round(r_squared25, 3)), size = 5, color = "black")

```

## Optimal dilution

**Error using 0.06 mL buffer**
```{r echo = FALSE}
# Print the results
cat("Mean Absolute Error (MAE) 0.06 mL:", round(mae06, 3), "\n")
cat("Standard Deviation of Residuals 0.06 mL:", round(std_dev06, 3), "\n")
```
**Error using 0.25 mL buffer**
```{r echo = FALSE}
cat("Mean Absolute Error (MAE) 0.25 mL:", round(mae25, 3), "\n") 
cat("Standard Deviation of Residuals 0.25 mL:", round(std_dev25, 3), "\n")

## Abs error and std dev. are larger for dilution 0.06 mL
```
From this we conclude that using **a 250 uL dilution provides more consistent results**

```{r echo = FALSE, include = FALSE}
# wflow_publish("./analysis/ELISA_Computation.Rmd")

```
