---
title: "Cortisol Concentration Values, Test4"
author: "Paloma Contreras"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

# Summary

**Cortisol value calculations** (includes bad quality samples, n = 41)


|                        | Min. | 1st Qu. | Median | Mean | 3rd Qu. | Max. | NA's |
|------------------------|------|---------|---------|---------|---------|-------|------|
|**A) Standard Method (mult. by sample dilution)**| 17.13 | 29.01  | 32.27  | 35.28 | 39.47 | 82.94 |    4  |
|**B) Spike-Corrected Method (Nist 2020)**| **-45.870** | -35.833 | -5.960 | -3.488 |  23.109 | **50.963**  |  4 |
|**C) Spike-Corrected (Sam's Method)**| **7.472** | 18.533    |24.559|  27.009| 31.196 |**80.804**    |   4 |


**Cortisol value calculations** (removed bad quality samples)

|                        | Min.   | 1st Qu. |  Median |  Mean   | 3rd Qu. | Max.   |  
|------------------------|--------|---------|---------|---------|---------|--------|
|**A) Standard Method (mult. by sample dilution)**| 18.27 | 29.27  | 31.54  | 34.13 | 37.73 | 69.17 |
|**B) Spike-Corrected Method (Nist 2020)**|**-39.371** | -34.855 | -20.577 | -14.325 |  -6.825 | **40.400** |
|**C) Spike-Corrected (Sam's Method)**| ** 11.80** | 18.45    |24.09|  24.05| 30.32 | **40.40**    |

**Results:**  

- Intra-assay CV: 14.5% 

- Intra-assay CV after removing low quality samples: 10%

- Inter-assay CV: **21%** (Bindings for 20mg sample diluted in 250 uL, no spike: 64.8% and 48% in test3 and test4, respectively)


**Conclusions**: 



**Concerns:**
Overall quality of the plate is not great, but serial dilusions show clear parallelism and standards have values within the expected

## Explanation of each variable used in calculations

- Ave_Conc_pg/ml: average ELISA reading per sample in pg/mL

- Weight_mg: hair weight in mg

- Buffer_nl: assay buffer volume in nL → we convert to mL

- Spike: binary indicator (1 = spiked sample)

- SpikeVol_uL: volume of spike added in µL

- Dilution: dilution factor (already present)

- Vol_in_well.tube_uL: total volume in well/tube in µL (for spike correction)

- std: standard reading value

- extraction: methanol volume ratio = vol added / vol recovered (e.g. 1/0.75 ml)


# Cortisol concentration calculations

```{r parameters}
# set reading value of spike (std1, 0.333 ug/dL), 
# and transforming to ug.dL

std <- (3191+3228)/2
std.r <- (std/10000)
std
std.r

# according to chatgpt, the spike's contribution is
# 1600 pg/mL, which is very similar to half of the reading for std 1 :]
```

```{r echo = FALSE, message=FALSE}
#path:
data_path = "./data/Test4"
library(knitr)
library(ggplot2)
library(broom)
library(paletteer)
library(dplyr)
```

Loading files and transforming units, *including low quality data*

```{r setup, include=TRUE, message=FALSE}
df <- read.csv(file.path(data_path,"Data_QC_flagged.csv"))
kable(tail(df)) 

# remove outlier
#df<- df[(df$Sample != "TP3A"),]

# Creating variables in indicated units
# dilution (buffer)
df$Buffer_ml <- c(df$Buffer_nl/1000)
df$Failed_samples[is.na(df$Failed_samples)] <- "OK"
table(df$Failed_samples)
df$Failed_samples[df$Failed_samples == "ABOVE 80% binding"] <- "Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;ABOVE 80% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV;UNDER 20% binding"] <- "High CV & Out of curve"
df$Failed_samples[df$Failed_samples == "HIGH CV"] <- "High CV"
df$Failed_samples[df$Failed_samples == "UNDER 20% binding"] <- "Out of curve"

df$Failed_samples <- factor(
  df$Failed_samples,
  levels = c(
    "OK",
    "Out of curve",
    "High CV",
    "High CV & Out of curve"
  )
)



# remove unnecessary information 
data <- df %>%
    dplyr::select(Wells, Sample, Category, Binding.Perc, Ave_Conc_pg.ml, Weight_mg, Buffer_ml, Spike, SpikeVol_uL, Dilution_sample, Dilution_spike, Extraction_ratio, Vol_in_well.tube_uL, Failed_samples) 



# remove duplicates
data <- data[!is.na(data$Binding.Perc), ]

kable(tail(data, 10))
dim(data)
```

## (A) Standard Calculation

Formula: 

((A/B) * (C/D) * E * 10,000) = F 

- A = μg/dl from assay output;
- B = weight (in mg) of hair subjected to extraction;
- C = vol. (in ml) of methanol added to the powdered hair;
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down;
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract;
- F = final value of hair CORT Concentration in pg/mg.


```{r A}
##################################
##### Calculate final values #####
##################################

# Transform to μg/dl from assay output
data$Ave_Conc_ug.dL <- c(data$Ave_Conc_pg.ml/10000)

data$Final_conc_pg.mg <- c(
    ((data$Ave_Conc_ug.dL) / data$Weight_mg) * # A/B *
      data$Extraction_ratio *                                  # C/D  *     
      data$Buffer_ml * 10000 * data$Dilution_sample)                 # E * 10000
data <- data[order(data$Sample),]
write.csv(data, file.path(data_path, "Data_cort_values_methodA.csv"), row.names = F)

# summary for all samples
data <- data %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL")) 
dim(data)
summary(data$Final_conc_pg.mg)

# summary for good quality samples only
temp <- data %>% 
  filter(Failed_samples == 
           "OK")
dim(temp)
summary(temp$Final_conc_pg.mg)


kable(tail(data, 7))


```

## (B) Accounting for Spike

We followed the procedure described in **Nist et al. 2020**:

"Thus, after pipetting 25μL of standards and samples into the appropriate wells of the 96-well assay plate, we added 25μL of the 0.333ug/dL standard to all samples,
resulting in a 1:2 dilution of samples. The remainder of the manufacturer’s protocol was
unchanged. We analyzed the assay plate in a Powerwave plate reader (BioTek, Winooski,
VT) at 450nm and subtracted background values from all assay wells. In the calculations, we
subtracted the 0.333ug/dL standard reading from the sample readings. Samples that resulted
in a **negative number were considered nondetectable**. We converted cortisol levels from
ug/dL, as measured by the assay, to pg/mg—based on the mass of hair collected and
analyzed using the following formula:

A/B * C/D * E * 10,000 * 2 = F

where 
- A = μg/dl from assay output; 
- B = weight (in mg) of collected hair; 
- C = vol. (in ml) of methanol added to the powdered hair; 
- D = vol. (in ml) of methanol recovered from the extract and subsequently dried down; 
- E = vol. (in ml) of assay buffer used to reconstitute the dried extract; 10,000 accounts for changes in metrics; 2 accounts for the dilution factor after addition of the spike; and 
- F = final value of hair cortisol concentration in pg/mg"


```{r B-calc}
dSpike <- data

##################################
##### Calculate final values #####
##################################

# spike is already divided by 10000 (unit is ug/dL)
dSpike$Final_conc_pg.mg <- 
  ifelse(
    dSpike$Spike == 1,    ## Only spiked samples
      ((dSpike$Ave_Conc_ug.dL - (std.r)) / # (A-spike) / B
        dSpike$Weight_mg) 
        * data$Extraction_ratio *                  # C / D
        dSpike$Buffer_ml * 10000 * 2,    # E * 10000 * 2
        dSpike$Final_conc_pg.mg  
    )


write.csv(dSpike, file.path(data_path, "Data_cort_values_methodB.csv"), row.names = F)

# summary for all samples
summary(dSpike$Final_conc_pg.mg)
dSpike$Sample
# summary for all samples
dSpike <- dSpike %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL")) 
dim(dSpike)
summary(dSpike$Final_conc_pg.mg)

# summary for good quality samples only
temp <- dSpike %>% 
  filter(Failed_samples == 
           "OK")
dim(temp)
summary(temp$Final_conc_pg.mg)

kable(tail(dSpike, 10))
```

## (C) Sam's calculation

- Simplify unnecessary unit transformations

- Account for spike considering dilution of both sample and the spike

Spike contribution (pg/mL) = (Vol. spike (mL) x Conc. spike (pg/mL) ) / Vol. reconstitution (mL) or total vol. in well (50uL) (depending on where the spike was added)

```{r C calc}
# Calculate contribution of spike according to the different volumes in which it was added
# Consider that contribution of spike in serial dilutions gets smaller 

# Vol. of spike transformed to mL
data$SpikeVol_ml <- data$SpikeVol_uL/1000
# Concentration of the spike:
std
# Vol. reconstitution (mL) is the total volume in tube or well (sample + spike), after adding spike.
# transform to mL
data$Vol_in_well.tube_ml <- data$Vol_in_well.tube_uL/1000

  ##( Spike vol. x Spike Conc.)
  ## ------------------------  / dilution = Spike contribution
  ##      Total vol. 
  
# Cortisol added by spike in wells: 0.0025 mL x 3200 pg/mL = 80 pg
# Calculate cort contribution of spike to each sample
data$Spike.cont_pg.mL <- ((data$SpikeVol_ml * std  / # Volume of spike * Spike concentration
                            data$Vol_in_well.tube_ml) / # divided by the total volume (spike + sample)
                              data$Dilution_spike) # resulting number changes depending on the dilution

dSpiked <- data
                          
##################################
##### Calculate final values #####
##################################


dSpiked$Final_conc_pg.mg <- 
     (((dSpiked$Ave_Conc_pg.ml - dSpiked$Spike.cont_pg.mL)) / # (A - spike) / B
      dSpiked$Weight_mg) *
     dSpiked$Extraction_ratio *      # C / D
      dSpiked$Buffer_ml * dSpiked$Dilution_sample    # E * 



write.csv(dSpiked, file.path(data_path, "Data_cort_values_methodC.csv"), row.names = F)

# summary for all samples

dSpiked <- dSpiked %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL")) 
dim(dSpiked)
summary(dSpiked$Final_conc_pg.mg)

# summary for good quality samples only
temp <- dSpiked %>% 
  filter(Failed_samples == 
           "OK")
dim(temp)
summary(temp$Final_conc_pg.mg)

kable(tail(dSpike, 10))

kable(head(dSpiked[!is.na(dSpiked$Final_conc_pg.mg) , c("Sample", "Final_conc_pg.mg", "Ave_Conc_pg.ml", "Spike.cont_pg.mL", "Binding.Perc", "Weight_mg", "Buffer_ml", "SpikeVol_uL", "Dilution_sample", "Dilution_spike", "Vol_in_well.tube_uL", "Extraction_ratio")],10))


```

# Plots 

## (A) Standard Calculation

```{r echo = FALSE, warning=FALSE }
# scatterplot method A

data <- data %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL")) #TA6","TA7","TB7","TC7", "TC6","TC5", #"TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1")

data$Spike <- replace(data$Spike, data$Spike == 1, 'Yes')
data$Spike <- replace(data$Spike, data$Spike == 0, 'No')
data$Buffer <- data$Buffer_ml
data$Buffer <- replace(data$Buffer, data$Buffer == 0.06, '60 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.11, '110 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.22, '220 uL')
data$Buffer <- replace(data$Buffer, data$Buffer == 0.25, '250 uL')


ggplot(data, aes(y = Final_conc_pg.mg, 
                 x = Binding.Perc, 
                 color = Failed_samples,
                 shape = Buffer)) +
  geom_point(size = 2.5, alpha = 0.85) +  
   geom_text(aes(label = Sample), size = 2.5, vjust = -0.65, hjust = -0.18) +
    theme_minimal() +  
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  #xlim(0,52) +
  labs(
    title = "(A) Standard Calculation Cortisol Values",
    y = "Final Concentration (pg/mg)",
    x = "Binding %") + 
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  ) + 
  scale_color_paletteer_d("vangogh::CafeTerrace")
```

## (B) Accounting for Spike
```{r echo = FALSE, warning=FALSE }

dSpike <- dSpike %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL"))
#TA6","TA7","TB7","TC7", "TC6","TC5", #"TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1")

dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 1, 'Yes')
dSpike$Spike <- replace(dSpike$Spike, dSpike$Spike == 0, 'No')
dSpike$Buffer <- dSpike$Buffer_ml
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.06, '60 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.11, '110 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.22, '220 uL')
dSpike$Buffer <- replace(dSpike$Buffer, dSpike$Buffer == 0.25, '250 uL')
head(dSpike)
# scatterplot
ggplot(dSpike, aes(y = Final_conc_pg.mg, 
                 x = Binding.Perc, 
                 color =  Failed_samples,
                 shape = Buffer)) +
  geom_point(size = 2.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
  #xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(B) Calculation Accounting for Spike",
    y = "Final Concentration (pg/mg)",
    x = "Binding %" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  ) + 
  scale_color_paletteer_d("vangogh::CafeTerrace")

```

## (C) Sam's calculation

```{r echo = FALSE, warning=FALSE }

dSpiked <- dSpiked %>%
  filter(!Sample %in% c("B0", "BE", "NSB", "POOL"))
#TA6","TA7","TB7","TC7", "TC6","TC5", #"TA2","TC3", "TD6", "TD7", "TB6", "TA1", "TB1", "TD1")
dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 1, 'Yes')
dSpiked$Spike <- replace(dSpiked$Spike, dSpiked$Spike == 0, 'No')
dSpiked$Buffer <- dSpiked$Buffer_ml
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.06, '60 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.11, '110 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.22, '220 uL')
dSpiked$Buffer <- replace(dSpiked$Buffer, dSpiked$Buffer == 0.25, '250 uL')

# scatterplot
dSpikedplot <- filter(dSpiked, Category == "A" | Category =="Standard" | Category == "B" | Category == "C" | Category == "D", !is.na(Binding.Perc)) 

ggplot(dSpikedplot, aes(y = Final_conc_pg.mg, 
                 x = Binding.Perc, 
                 color = Failed_samples,
                 shape = Buffer)) +
  geom_point(size = 2.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
 #   ylim(-26,30) +
  #xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(C) Sam's calculation",
    y = "Final Concentration (pg/mg)",
    x = "Binding %" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  )+ 
  scale_color_paletteer_d("vangogh::CafeTerrace")


dSpikedplot <- filter(dSpiked, Category == "P", !is.na(Binding.Perc)) 

ggplot(dSpikedplot, aes(y = Final_conc_pg.mg, 
                 x = Binding.Perc, 
                 color = Failed_samples,
                 shape = Buffer)) +
  geom_point(size = 2.5,  alpha = 0.85) +  
  geom_text(aes(label = Sample), size = 3, vjust = -1, hjust = -0.1) +
  theme_minimal() +  
 #   ylim(-26,30) +
  #xlim(0,52) +
  geom_hline(yintercept = 0, 
             linetype = "dashed", color = "red") +
  labs(
    title = "(C) Sam's calculation, Precision",
    y = "Final Concentration (pg/mg)",
    x = "Binding %" ) +
  theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 17, face = "bold"),
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 10)  
  )+ 
  scale_color_paletteer_d("vangogh::CafeTerrace")

```



# Evaluation method C


```{r}

ggplot(dSpiked, aes(y = Final_conc_pg.mg, 
                  x = Spike.cont_pg.mL, 
                  color = (Failed_samples))) + 
                 # fill = factor(Spike.cont_pg.mL))) +
  geom_point(size = 2.5) +  
  geom_text(label = c(dSpiked$Sample), nudge_y = 0.95, nudge_x = -1.2) +
 # geom_hline(yintercept = mean(data$Final_conc_pg.mg), 
  #           color = "gray80",
   #          linetype = "dashed") +
  theme_minimal() +  
  labs(
    title = "Final Cort Concentration and Contribution of spike",
    y = "Final Concentration (pg/mg)",
    x = "Contribution of spike (pg/ml)"  
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.title = element_text(size = 14),  
    axis.text = element_text(size = 12) 
  ) 

```

