---
title: "ELISA_Simulation_test4"
output: html_document
date: "2025-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
```

# Introduction
Here I compare two calculation methods for determining final cortisol concentrations (in pg/mg) from ELISA values of spiked samples

Nist Method: Assumes subtraction of an averaged spike concentration converted to the same unit (ug/dL), not accounting for the actual spike volume per sample.

Sam Method: Method that subtracts the actual contribution of the spike (based on its volume and concentration) before calculating the final concentration.


# Simulate dataset
```{r}

# Simulate toy dataset
set.seed(42)
n <- 35

dSpike <- data.frame(
  SampleID = paste0("S", 1:n),
  Ave_Conc_pg.ml = runif(n, 33, 15600),  # Simulated ELISA values
  Buffer_ml = rep(0.25, n),            # Constant assay buffer volume (E)
  Spike = rep(1, n)                     # All samples are spiked
)

dSpike$Ave_Conc_ug.dL <- c(dSpike$Ave_Conc_pg.ml/10000) # transform to microliter/deciliter
dSpike$Weight_mg <- 50              # Hair weights
extraction <- 1 / 0.75    # Methanol extraction ratio (in mL)

# Two versions of the calculation
std <- (3191+3228)/2
std_ug.dL  <- (std/10000)
std_ug.dL_Nist <- std_ug.dL   # Nist
std_pg.ml_Sam <- std  # Sam 

# Nist calculation 
# A/B * C/D * E * 10,000 * 2 = F

dSpike$pg.mg_Nist <- ((dSpike$Ave_Conc_ug.dL - std_ug.dL_Nist) / dSpike$Weight_mg) * # A/B *
  extraction *              # C/D *
  dSpike$Buffer_ml * 10000 * 2  # E * 10000

summary(dSpike$pg.mg_Nist)


# Sam (Accounting for spike contribution)
##  ( Spike vol. x Spike Conc.)
##  ---------------------------  / dilution = Spike contribution
##          Total vol. 
  
SpikeVol_uL <- 25 # in microliters
SpikeVol_ml <- SpikeVol_uL/1000  # transform to mL
dilution <- 1
Vol_in_well.tube_ml <- 0.05   # in mL

# calculate contribution of spike
dSpike$Spike.cont_pg.mL <- (((SpikeVol_ml * std_pg.ml_Sam  ) / # Volume of spike * Spike concentration
                            Vol_in_well.tube_ml) / # divided by the total volume (spike + sample)
                              dilution) # resulting number changes depending on the dilution
# Calculating final vals
dSpike$pg.mg_Sam <- 
      ((dSpike$Ave_Conc_pg.ml - dSpike$Spike.cont_pg.mL) / # (A - spike) / B
      dSpike$Weight_mg) *
      extraction *      # C / D
      dSpike$Buffer_ml    # E * 

# Compare in a tidy format


dLong <- pivot_longer(
  dSpike,
  cols = c("pg.mg_Nist", "pg.mg_Sam"),
  names_to = "Method",
  values_to = "Final_pg.mg"
)

# Plot
ggplot(dLong, aes(x = SampleID, y = Final_pg.mg, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Comparison of Hair Cortisol (pg/mg)",
    subtitle = "Nist vs Sam Spike Subtraction",
    y = "Cortisol (pg/mg)", x = "Sample"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("pg.mg_Nist" = "red", "pg.mg_Sam" = "darkgreen"),
                    labels = c("Nist", "Sam"))

dWide <- dSpike[, c("pg.mg_Nist", "pg.mg_Sam")]
summary(dWide)
mean(abs(dWide$pg.mg_Nist - dWide$pg.mg_Sam))

```
