---
title: "Test5_design"
output: html_document
date: "2025-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# VALIDATION 1: Spike-Recovery Test

🎯 **Purpose**:
To assess how much known cortisol is recovered from a pulverized extract. This helps detect matrix interference (e.g., suppression or enhancement).

🔬 **Materials**:

- Pulverized, washed blank hair sample (very low or no cortisol)
- Cortisol standard solution (e.g., 1000 pg/mL in assay buffer)
- ELISA plate with standard curve

🧪 **Procedure**:

| Tube |	Contents |	Description | Explanation |
-------|---------|--------|-------|
A	| Pulverized hair + methanol (extraction only) | Unspiked control | get a baseline value
B	| Pulverized hair + methanol + known cortisol (e.g., 100 pg) |	Spiked sample | checks if you recover what you added
C	| Cortisol in methanol (no hair) |	Solvent control | ensures no cortisol loss or pipetting error

**REMEMBER**: Let the spiked tube sit for 30–60 min to simulate binding, then proceed with extraction and ELISA as usual.

Reconstitute the dried extract and run both A and B in duplicate on the plate.

🧮 **Calculation**:

Recovery (%) = ((Measured pg − Background pg (A)) /  Spiked pg ) * 100

Expected:

80–120% recovery is generally considered acceptable.
If <80% or >120%, matrix effects are likely.

# 🧪 VALIDATION 2: Parallelism Test
🎯 **Purpose**:

To determine whether a serial dilution of your extract runs parallel to the standard curve, indicating minimal matrix effects (no inhibition or enhancement).

🔬 **Materials:**

- A high-concentration pulverized extract (from a real or pooled sample)
- Your ELISA standard curve
- Assay buffer for dilution

🧪 **Plate Layout** :

Well |	Dilution |	Contents|
|----|----|----|
S1 |	Neat extract|	100%
S2	|1:2 dilution|	50%
S3	|1:4 dilution	|25%
S4	|1:8 dilution	|12.5%
S5	|1:16 dilution	|6.25%
...	...	

Run all diluted samples in duplicate.

📉 **Plotting & Interpreting**:

Plot absorbance (y-axis) vs. log concentration (x-axis) for:

- The standard curve
- Your diluted extract
- Check for parallelism between the sample and the standard curve.

Expected: 

If sample curve is parallel to the standard, no major matrix effect is present.
If the sample curve bends or flattens relative to the standard, there's likely matrix interference.

🧪 **Summary of Both Tests**

|Test |	What It Tells You	| Outcome to Look For|
|---|---|---|
|Spike-Recovery|	Do matrix components suppress/enhance cortisol signal?|	Recovery 80–120%|
|Parallelism|	Do serial dilutions track with the standard curve?|	Same shape/parallel curves|





```{r}

```

